---
title: "Stand_Stock_Tables_MarkDown"
output: html_document
date: "2024-03-14"
---

#load necessary packages

```{r}
library(plyr)
library(stringr)
library(magrittr)
library(dplyr)
library(tidyverse)
library(cffdrs)
library(dplyr)
library(magrittr)
library(XLConnect)
library(openair)
library(plyr)
library(ggplot2)
library(grid)
library(tidyr)
library(reshape2)
library(readxl)
library(testthat)
library(GA)
library(ftsa)
library(xlsx)
library(devtools)
library(firebehavioR)
library(kableExtra)
library(webshot2)
library(magick)
library(pagedown)
```

## Set File Paths change this everytime

```{r}
# set file path directory
path = "Z:/Projects/MOF/NE WRR Prescriptions 2025/Modelling/TR_LionsBurn"
out = path
name = "TR_LionsBurn"
project=name


# Prefix for SNAP summary file names
snap_prefix = "/SNAP/"
cath_prefix = "/Cut Specs for run/"
s_s_prefix = "/Stand_StockTables/"


# Specify the path to the Excel file
excel_file <- paste0(path, cath_prefix, "Cathro.xlsx")
```

#1.Export Files

  #Export the Cathro by selecting the "Report" and "Cathro" Button in SNAP. Then select "EXPORT" from the top left corner and save as "Excel (2007+)". The file will open on your computer and save it inside the SNAP folder inside your projects modelling folder as "Cathro.xlsx".

  #read in Extra export csv to establish the number and names of treatments:
  go to snap and export SNAP_ex.csv document by loading report for your project of interest, then clicking Analysis
  then selecting your project and choosing "Extra" from the check box list on the left. Make sure some variables 
  are selected then click okay. Then in the top right corner of the document that pops up click export and save file 
  The file will open up automatically in excel. Navigate to the top corner and manually resave it as a csv as [Project Name]_EXTRA.csv.
  Redo this for both the "Overstory" and "Understory", and "Fuels" check boxes, saving them as "[Project Name]_OS.csv" and "[Project Name]_US.csv", and [Project Name]_FUELS.csv respectively.

Make folders inside Modeling Folder: 1 for each treatment
Stand_StockTables
  TU1_tables
  TU2_tables
  TU3_tables
  TU4_tables
  ...
  ..
  .
  
#2.Load Files & Create Templates and CBH
This section will load in your exported CSVs and create folders for your stand and stock tables:
  -Will also create templates for your key files:
  OS_BA, OS_Vol, OS_SPH, US_SPH, and cbh and ht files
  -then you fill them in!

```{r}
# Load in SNAP summaries and cathro
Snap_OS = read.csv(paste0(path,snap_prefix,name,"_OS.csv"))
Snap_US = read.csv(paste0(path,snap_prefix,name,"_US.csv"))
Snap_EX= read.csv(paste0(path,snap_prefix,name,"_EXTRA.csv"))
n_tre<-unique(Snap_EX$Stratum)
folder_paths<-(paste0(path,s_s_prefix))

#decide if L1 (12-17) is in overstory or not
L1_OS<-c(FALSE,FALSE)

#OS Temp
OSTEMP<-data.frame(
  DBH.Class= c("12.5 - 17.5","17.5 - 22.5","22.5 - 27.5","27.5 - 35","35 - 45","45+"),
  Total= c(rep(0,6))
)
USTEMP<-data.frame(
  Layer= c("Layer 4 (<1.3m)","Layer 3 (2.5-7.49)","Layer 2 (7.5-12.49)","Layer 1 (12.5-17.5)"),
  Total= c(rep(0,4))
)
#cut specs
cut_spec<-data.frame(
    Stand.Layer=c("L4","L3","L2","L1","L1","L1","L1","L1","L1"),
    DBH.Class=c("0 - 1.5","1.5 - 7.5","7.5 - 12.5","12.5 - 17.5","17.5 - 22.5" ,"22.5 - 27.5","27.5 - 35","35 - 45", "45+")
  )

for (treatment in n_tre) {
  
  # Create folder names based on Stratum values
  treatment_folder <- paste0(folder_paths,treatment,"_tables")
  
  # Create the directory if it doesn't already exist
  if (!dir.exists(treatment_folder)) {
    dir.create(treatment_folder, recursive = TRUE)
  }
  
  #
  index <- which(n_tre == treatment)
  
  if(!L1_OS[index]){
  OSOUT<-OSTEMP[-1,]
  USOUT<-USTEMP
  }else{
  OSOUT<-OSTEMP
  USOUT<-USTEMP
  }
 
  #load snap
  st_os<-Snap_OS %>% dplyr::filter(Stratum == treatment)
  st_us<-Snap_US %>% dplyr::filter(Stratum == treatment)
  
  #create BA, Vol, SPH, CBH files
  n_spp_OS<- unique(st_os$Spp)
  n_spp_US<-unique(st_us$SPP)
  
  #add columns
  OSOUT[n_spp_OS]<-0
  USOUT[n_spp_US]<-0
  
  #EXPORT
  OSOUT<- OSOUT %>% dplyr::select(-Total)
  USOUT<- USOUT %>% dplyr::select(-Total)
  
    #
  write.csv(OSOUT,paste0(treatment_folder,"/OS_SPH.csv"),row.names=FALSE)
  write.csv(OSOUT,paste0(treatment_folder,"/OS_BA.csv"),row.names=FALSE)
  write.csv(OSOUT,paste0(treatment_folder,"/OS_Vol.csv"),row.names=FALSE)
  write.csv(USOUT,paste0(treatment_folder,"/US_SPH.csv"),row.names=FALSE)
  
  #fake cutspec
  all_sp<-unique(c(n_spp_OS,n_spp_US))
  all_sp <- ifelse(all_sp %in% c("Dead", "DU"), "DP", all_sp)
  cut_spec[all_sp]<-0
  write.csv(cut_spec,paste0(treatment_folder,"/cuttingSpecs_",treatment,".csv"),row.names=FALSE)


  #Ht and cbh
  #OSht_names<- paste0(n_spp_OS,"_Ht")
  #OScbh_names<-paste0(n_spp_OS,"_CBH")
  #OS_ALL<-c(OSht_names,OScbh_names)
  #OSHT<-OSOUT %>% dplyr::select(-n_spp_OS)
  #OSHT[OS_ALL]<-0
  
  #generate ht and cbh   DBH.Class= c("17.5 - 22.5","22.5 - 27.5","27.5 - 35","35 - 45","45+"),

  #create cbh label
  st_os$dbh_label <- ifelse(st_os$DBH >= 12.5 & st_os$DBH < 17.5, "12.5 - 17.5",
    ifelse(st_os$DBH >= 17.5 & st_os$DBH < 22.5, "17.5 - 22.5",
                     ifelse(st_os$DBH >= 22.5 & st_os$DBH < 27.5, "22.5 - 27.5",
                     ifelse(st_os$DBH >= 27.5 & st_os$DBH < 35, "27.5 - 35",
                     ifelse(st_os$DBH >= 35 & st_os$DBH < 45, "35 - 45", "45+")))))
  
  cbh_os<- st_os %>%
    dplyr::group_by(dbh_label,Spp) %>% 
    dplyr::summarise(
      cbh=round(mean(CBH..0.1m.,na.rm=T),2)
    )
  ht_os<- st_os %>%
    dplyr::group_by(dbh_label,Spp) %>% 
    dplyr::summarise(
      ht=round(mean(Total.Height..m.,na.rm=T),2)
    )
  ht_short<-pivot_wider(ht_os,names_from = "Spp", values_from = "ht",names_glue = "{Spp}.Ht")
  cbh_short<-pivot_wider(cbh_os,names_from = "Spp", values_from = "cbh",names_glue = "{Spp}.CBH")
  OSHTCBH<-left_join(ht_short,cbh_short, by="dbh_label")
  colnames(OSHTCBH)[1]<-"DBH.Class"
  
  #make sure all layers present
  expected_dbh <- OSTEMP$DBH.Class
  OSHTCBH <- OSTEMP %>%
    dplyr::select(DBH.Class) %>%
    left_join(OSHTCBH, by = "DBH.Class")
  
  #export
  write.csv(OSHTCBH,paste0(treatment_folder,"/OS_Ht_CBH.csv"),row.names=FALSE)

  #Layer= c("Layer 4 (<1.3m)","Layer 3 (2.5-7.49)","Layer 2 (7.5-12.49)","Layer 1 (12.5-17.5)")
  #USht_names<- paste0(n_spp_US,"_Ht")
  #UScbh_names<-paste0(n_spp_US,"_CBH")
  #US_ALL<-c(USht_names,UScbh_names)
  #USHT<-USOUT %>% dplyr::select(-n_spp_US)
  #USHT[US_ALL]<-0
  
  
  cbh_us<- st_us %>%
    dplyr::group_by(Layer,SPP) %>% 
    dplyr::summarise(
      cbh=round(mean(CBH..0.1m.,na.rm=T),2)
    )
  ht_us<- st_us %>%
    dplyr::group_by(Layer,SPP) %>% 
    dplyr::summarise(
      ht=round(mean(Height..0.1m.,na.rm=T),2)
    )
  ht_short_us<-pivot_wider(ht_us,names_from = "SPP", values_from = "ht",names_glue = "{SPP}.Ht")
  cbh_short_us<-pivot_wider(cbh_us,names_from = "SPP", values_from = "cbh",names_glue = "{SPP}.CBH")
  USHTCBH<-left_join(ht_short_us,cbh_short_us, by="Layer")
  colnames(USHTCBH)[1]<-"Layer"

  #make sure all layers present
  expected_layers <- USTEMP$Layer
  USHTCBH <- USTEMP %>%
    dplyr::select(Layer) %>%
    left_join(USHTCBH, by = "Layer")
  
  #export
  write.csv(USHTCBH,paste0(treatment_folder,"/US_Ht_CBH.csv"),row.names=FALSE)
}


```
  
#3. Manual US Changes
  
```{r}
# Assign height values based on Layer if Height is NA
Snap_US$Height..0.1m.[is.na(Snap_US$Height..0.1m.) & Snap_US$Layer == "Layer 2 (7.5-12.49)"] <- 5
Snap_US$Height..0.1m.[is.na(Snap_US$Height..0.1m.) & Snap_US$Layer == "Layer 3 (2.5-7.49)"] <- 3
Snap_US$Height..0.1m.[is.na(Snap_US$Height..0.1m.)] <- 1  
# Assign CBH values based on Layer if CBH is NA
Snap_US$CBH..0.1m.[is.na(Snap_US$CBH..0.1m.) & Snap_US$Layer == "Layer 2 (7.5-12.49)"] <- 1.5
Snap_US$CBH..0.1m.[is.na(Snap_US$CBH..0.1m.) & Snap_US$Layer == "Layer 3 (2.5-7.49)"] <- 1
Snap_US$CBH..0.1m.[is.na(Snap_US$CBH..0.1m.)] <- 0  # default for others


```  

#4.Strata Data: Vol, BA, SPH
  
  4.1 Create Vol BA, SPH from exported Snap documents:
  
```{r}
#read in master xls
#test_cath<-read.xlsx("C:/Users/Holden/OneDrive/Desktop/test.xls",1)
Cath_master<-read.xlsx(excel_file,1)
colnames(Cath_master)[1]<-c("Focus_Column")

No_Trees_df<-data.frame(
  DBH.Class= c("17.5 - 22.5","22.5 - 27.5","27.5 - 35","35 - 45","45+", "Total"),
  Total= c(rep(0,6))
)
No_Trees_US<-data.frame(
  Layer= c("Layer 4 (<1.3m)","Layer 3 (2.5-7.49)","Layer 2 (7.5-12.49)","Layer 1 (12.5-17.5)"),
  Total= c(rep(0,4))
)

#import stratas
stratas<-unique(Snap_EX$Stratum)


#Iterate over treatments
for(i in 1:length(stratas)){
  st<-stratas[i]
#
  treat_OS <- Snap_OS %>%
  filter(Stratum == st)
  treat_US <- Snap_US %>%
  filter(Stratum == st)
  
  #get number of species then add 2 for DU and DP
  n_spp<-length(unique(treat_OS$Spp))
  
 if (st %in% treat_OS$Stratum) {
#extract overstory Basal Area Summary
BA<- "Overstory Basal Area Summary"
#find rows where overstory BA starts: should refer to all tratments
rows_start_BA <- Cath_master[Cath_master$Focus_Column == BA, ]
#get row numbers
row_numbers_BA <- which(rownames(Cath_master) %in% rownames(rows_start_BA))
#get specific row number for start of exportable part by iterating down list of row numbers you extracted above
BA_start<-row_numbers_BA[i]+2
#add end row number based on 1 header row and 6 dbh class rows(including total)
BA_end<-row_numbers_BA[i]+8
#find number of columnns for SPH df part of Cath report: take number of species and add 6. 5 of which are for the merged columns at the beginning, and 1 of which are for the total column.
if(n_spp == 0){   col_end_BA<- n_spp+5 } else{   col_end_BA<-n_spp+7 }
#read in specific DF with row indices
BA_xls<-xlsx::read.xlsx(excel_file,1,startRow =BA_start,endRow =BA_end,colIndex =c(5:col_end_BA))
BA_xls[is.na(BA_xls)] <- 0

if(colnames(BA_xls)[1] != "DBH.Class"){
  colnames(BA_xls)<-BA_xls[1,]
  colnames(BA_xls)[1]<-"DBH.Class"
}else{
  colnames(BA_xls)[1]<-"DBH.Class"
}

if(BA_xls$DBH.Class[1]=="DBH Class" | BA_xls$DBH.Class[1]=="DBH.Class" ){
  BA_xls<-BA_xls[-1,]
}

write.csv(BA_xls,paste0(path,s_s_prefix,st,"_tables/","OS_BA.csv"),row.names = FALSE)
 } else{
  BA_xls<-No_Trees_df
  write.csv(BA_xls,paste0(path,s_s_prefix,st,"_tables/","OS_BA.csv"),row.names = FALSE)

}

  if (st %in% treat_OS$Stratum) {
#extract overstory Volume Summary
Vol<- "Overstory Volume Summary"
#find rows where overstory Vol starts: should refer to all tratments
rows_start_Vol <- Cath_master[Cath_master$Focus_Column == Vol, ]
#get row numbers
row_numbers_Vol <- which(rownames(Cath_master) %in% rownames(rows_start_Vol))
#get specific row number for start of exportable part by iterating down list of row numbers you extracted above
Vol_start<-row_numbers_Vol[i]+2
#add end row number based on 1 header row and 6 dbh class rows(including total)
Vol_end<-row_numbers_Vol[i]+8
#find number of columns for SPH df part of Cathro report: take number of species and add 6. 5 of which are for the merged columns at the beginning, and 1 of which are for the total column.
if(n_spp == 0){   col_end_Vol<- n_spp+8 } else{   col_end_Vol<-n_spp+10 }
#read in specific DF with row indices
Vol_xls<-xlsx::read.xlsx(excel_file,1,startRow =Vol_start,endRow =Vol_end,colIndex =c(5:col_end_Vol))
#remove_NA's
Vol_xls[is.na(Vol_xls)]<-0

if(colnames(Vol_xls)[1] != "DBH.Class"){
  colnames(Vol_xls)<-Vol_xls[1,]
  colnames(Vol_xls)[1]<-"DBH.Class"
}else{
  colnames(Vol_xls)[1]<-"DBH.Class"
}

if(Vol_xls$DBH.Class[1]=="DBH Class" | Vol_xls$DBH.Class[1]=="DBH.Class" ){
  Vol_xls<-Vol_xls[-1,]
}

#export CSVs
write.csv(Vol_xls,paste0(path,s_s_prefix,st,"_tables/","OS_Vol.csv"),row.names = FALSE)
} else{
  Vol_xls<-No_Trees_df
  write.csv(Vol_xls,paste0(path,s_s_prefix,st,"_tables/","OS_Vol.csv"),row.names = FALSE)

}
  
  if (st %in% treat_OS$Stratum) {
#extract overstory SPH Summary
SPH<- "Overstory SPH Summary"
#find rows where overstory SPH starts: should refer to all tratments
rows_start_SPH <- Cath_master[Cath_master$Focus_Column == SPH, ]
#get row numbers
row_numbers_SPH <- which(rownames(Cath_master) %in% rownames(rows_start_SPH))
#get specific row number for start of exportable part by iterating down list of row numbers you extracted above
SPH_start<-row_numbers_SPH[i]+2
#add end row number based on 1 header row and 6 dbh class rows(including total)
SPH_end<-row_numbers_SPH[i]+8
#find number of columns for SPH df part of Cathro report: take number of species and add 6. 5 of which are for the merged columns at the beginning, and 1 of which are for the total column.
if(n_spp == 0){   col_end_SPH <- n_spp + 8 } else{   col_end_SPH<-n_spp+10 }
#read in specific DF with row indices
SPH_xls<-xlsx::read.xlsx(excel_file,1,startRow =SPH_start,endRow =SPH_end,colIndex =c(5:col_end_SPH))
#remove_NA's
SPH_xls[is.na(SPH_xls)]<-0

if(colnames(SPH_xls)[1] != "DBH.Class"){
  colnames(SPH_xls)<-SPH_xls[1,]
  colnames(SPH_xls)[1]<-"DBH.Class"
}else{
  colnames(SPH_xls)[1]<-"DBH.Class"
}
if(SPH_xls$DBH.Class[1]=="DBH.Class" | SPH_xls$DBH.Class[1]=="DBH Class" ){
  SPH_xls<-SPH_xls[-1,]
}

#export CSVs
write.csv(SPH_xls,paste0(path,s_s_prefix,st,"_tables/","OS_SPH.csv"),row.names = FALSE)
} else{
  SPH_xls<-No_Trees_df
  write.csv(SPH_xls,paste0(path,s_s_prefix,st,"_tables/","OS_SPH.csv"),row.names = FALSE)

}
  
  
   if (st %in% treat_US$Stratum) {
     
#extract understory SPH Summary
sph<- "Stratum UNDERSTORY - SPH"
#find rows where understory SPH starts: should refer to all tratments
rows_start_sph <- Cath_master[Cath_master$Focus_Column == sph, ]
#get row numbers
row_numbers_sph <- which(rownames(Cath_master) %in% rownames(rows_start_sph))
#get specific row number for start of exportable part by iterating down list of row numbers you extracted above
sph_start<-row_numbers_sph[i]+2
#add end row number based on 1 header row and 6 dbh class rows(including total)
n_layer<- length(unique(treat_US$Layer))
if(n_layer == 3){   sph_end<- row_numbers_sph[i]+n_layer+2 } else{   sph_end<-row_numbers_sph[i]+n_layer+2 }

#find number of columns for SPH df part of Cathro report: take number of species and add 6. 5 of which are for the merged columns at the beginning, and 1 of which are for the total column.
if(n_spp == 0){   col_end_sph<- n_spp+8 } else{   col_end_sph<-n_spp+10 }
#read in specific DF with row indices
sph_xls<-xlsx::read.xlsx(excel_file,1,startRow =sph_start,endRow =sph_end,colIndex =c(5:col_end_sph))
#remove_NA's
sph_xls[is.na(sph_xls)]<-0

if(colnames(sph_xls)[1] != "Layer"){
  colnames(sph_xls)<-sph_xls[1,]
  colnames(sph_xls)[1]<-"Layer"
}else{
  colnames(sph_xls)[1]<-"Layer"
}

if(sph_xls$Layer[1]=="Layer"| sph_xls$Layer[1]=="Layer" ){
  sph_xls<-sph_xls[-1,]
}
#export CSVs
write.csv(sph_xls,paste0(path,s_s_prefix,st,"_tables/","US_SPH.csv"),row.names = FALSE)
} else{
  sph_xls<-No_Trees_df
  write.csv(sph_xls,paste0(path,s_s_prefix,st,"_tables/","US_SPH.csv"),row.names = FALSE)
}
}

```

  4.2 Create CBH, HT, SPH files from exported Snap documents:

```{r}
# Assuming Snap_OS is your dataframe add DBH class column
Snap_OS$DBH.Class <- ifelse(Snap_OS$DBH >= 12.5 & Snap_OS$DBH < 17.5, "12.5 - 17.5",
                            ifelse(Snap_OS$DBH >= 17.5 & Snap_OS$DBH < 22.5, "17.5 - 22.5",
                            ifelse(Snap_OS$DBH >= 22.5 & Snap_OS$DBH < 27.5, "22.5 - 27.5",
                                   ifelse(Snap_OS$DBH >= 27.5 & Snap_OS$DBH < 35, "27.5 - 35",
                                          ifelse(Snap_OS$DBH >= 35 & Snap_OS$DBH < 45, "35 - 45",
                                                 "45+")))))

DBH_order<- c("12.5 - 17.5","17.5 - 22.5","22.5 - 27.5","27.5 - 35","35 - 45", "45+")

#create SPH, CBH and Ht dataframes for OS and US
stratas<- unique(Snap_EX$Stratum)

for (i in 1:length(stratas)) {
  st<-stratas[i]
  #
  treat_OS <- Snap_OS %>%
  filter(Stratum == st)
  treat_US <- Snap_US %>%
  filter(Stratum == st)
  
###########################################
#######################################################    
  #Get Mean Height and CBH for OS
if (st %in% treat_OS$Stratum) {
 OS_ht_cbh <- treat_OS %>%
  filter(!is.na(DBH)) %>%
  dplyr::group_by(Spp,DBH.Class) %>%
  dplyr::summarise(.Ht = mean(Total.Height..m.),
            .CBH = mean(CBH..0.1m., na.rm = TRUE)) %>% 
   tidyr::pivot_wider(names_from = Spp, values_from = c(.Ht, .CBH), names_glue = "{Spp}{.value}")
 OS_ht_cbh[, -1] <- lapply(OS_ht_cbh[, -1], as.numeric)
 #
 OS_ht_cbh$DBH.Class <- factor(OS_ht_cbh$DBH.Class, levels = DBH_order)
 
# Identify missing levels
missing_levels <- setdiff(DBH_order, unique(OS_ht_cbh$DBH.Class))

# Create a data frame of missing levels with NAs for other columns
if(!is.na(missing_levels[1])){
missing_df <- data.frame(DBH.Class = factor(missing_levels, levels = DBH_order))
for(col in setdiff(names(OS_ht_cbh), "DBH.Class")) {
  missing_df[[col]] <- NA
}
}
missing_df <- data.frame(DBH.Class = factor(missing_levels, levels = DBH_order))
OS_ht_cbh <- dplyr::bind_rows(OS_ht_cbh, missing_df)
 OS_ht_cbh <- arrange(OS_ht_cbh, DBH.Class)
  #
 OS_ht_cbh[is.na(OS_ht_cbh)]<-0
 write.csv(OS_ht_cbh,paste0(path,s_s_prefix,st,"_tables/","OS_Ht_CBH.csv"),row.names = FALSE)
} else {
  OS_ht_cbh <- No_Trees_df
 write.csv(OS_ht_cbh,paste0(path,s_s_prefix,st,"_tables/","OS_Ht_CBH.csv"),row.names = FALSE)
}
###########################################################################   
   #Get Mean Height and CBH for US
 if (st %in% treat_US$Stratum) {
 US_ht_cbh <- treat_US %>%
  filter(!is.na(Height..0.1m.)) %>%
  dplyr::group_by(SPP, Layer) %>%
  dplyr::summarise(.Ht = mean(Height..0.1m.),
            .CBH = weighted.mean(CBH..0.1m.,X..of.Trees, na.rm = TRUE)) %>%
  tidyr::pivot_wider(names_from = SPP, values_from = c(.Ht, .CBH), names_glue = "{SPP}{.value}")
  US_ht_cbh[, -1] <- lapply(US_ht_cbh[, -1], as.numeric)
US_ht_cbh[is.na(US_ht_cbh)]<-0
 write.csv(US_ht_cbh,paste0(path,s_s_prefix,st,"_tables/","US_Ht_CBH.csv"),row.names = FALSE)
} else {
  US_ht_cbh <- No_Trees_US
   write.csv(US_ht_cbh,paste0(path,s_s_prefix,st,"_tables/","US_SPH.csv"),row.names = FALSE)
}

}


``` 
 

#5.Create Manual Cutting Specs Files
  
  Find Template Here: "Z:\Scripts\FronteraCodez\Modeling Templates\Modeling\cuttingSpecs.csv"
  
  Modify and delete Species as needed and save it like below:
  
  Document should be called cuttingSpecs in Stand_StockTables/Tu"treatment number"_tables folder.
------
  ie for treatment 1 cuttingSpecs.csv document is located at
  
  ~(Project) Modeling/Stand_StockTables/Tu1_Tables/cuttingSpecs.csv
  
  and one called
  
  cuttingSpecs_[trname].csv
-------

  Table must have these two columns first: 
Stand.Layer
L4
L3
L2
L1
L1
L1
L1
L1
L1

-----------------------
DBH.Class
0-1.5
1.5-7.5
7.5-12.5
12.5-17.5
17.5-22.5
22.5-27.5
27.5-35
35-45
45+

Then a column for each appropriate species labeled with code and first letter capitalized. Non-changes should be zero and values should be in percent. (ie 90% --> 90, 75% -->75)
  

#6. Create Stand and Stock Tables
  
  6.1 Run code below for each treatment unit should produce stand and stock tables in appropriate folder as html and pdf and excel in the output folder

```{r}
#import stratas
stratas<-unique(Snap_EX$Stratum)

#iterate over treatments
for(i in 1:length(unique(Snap_EX$Stratum))){
 
TR<- stratas[i]
table_folder<- paste0(TR,"_tables/")

###########################################################################################################
cuttingSpecs <- read.csv(file = paste0(path,s_s_prefix,table_folder,"cuttingSpecs_",TR,".csv"), row.names = NULL, stringsAsFactors = FALSE)
OS_SPH <- read.csv(file = paste0(path,s_s_prefix,table_folder,"OS_SPH.csv"), row.names = NULL, stringsAsFactors = FALSE)
OS_BA <- read.csv(file = paste0(path,s_s_prefix,table_folder,"OS_BA.csv"), row.names = NULL, stringsAsFactors = FALSE)
OS_Vol <- read.csv(file = paste0(path,s_s_prefix,table_folder,"OS_Vol.csv"), row.names = NULL, stringsAsFactors = FALSE)
OS_Ht_CBH <- read.csv(file = paste0(path,s_s_prefix,table_folder,"OS_Ht_CBH.csv"), row.names = NULL, stringsAsFactors = FALSE)
US_SPH <- read.csv(file = paste0(path,s_s_prefix,table_folder,"US_SPH.csv"), row.names = NULL, stringsAsFactors = FALSE)
US_Ht_CBH <- read.csv(file = paste0(path,s_s_prefix,table_folder,"US_Ht_CBH.csv"), row.names = NULL, stringsAsFactors = FALSE)
###remove first column
#remove columns that start with NA
US_SPH <- US_SPH[, !grepl("^NA", names(US_SPH))]
OS_SPH<- OS_SPH[, !grepl("^NA", names(OS_SPH))]

#######################
coniferList <- c("Ba","Bl", "Bg","Bb","Cw", "Fd", "Hw","T","Pl", "Sx","Sb","Sw", "Lw", "Lt", "Pw","Yc", "Fdi", "Fdc", "Py")
nonConiferList <- c("Act","Acb","Ac","At","Ep","DP", "DU","Dead", "Dr")

# 1. Overstorey

# Height and CBH

OS_Ht_CBH_melt <- reshape2::melt(OS_Ht_CBH, id = "DBH.Class")
OS_Ht_CBH_melt$Species <- gsub("\\.", "", substr(OS_Ht_CBH_melt$variable, start = 1, stop = 3))

#check for yew
OS_Ht_CBH_melt$Species[OS_Ht_CBH_melt$Species == "T."] <- "T"

OS_Ht <- OS_Ht_CBH_melt[grep(pattern = "Ht", OS_Ht_CBH_melt$variable, fixed = TRUE),]
names(OS_Ht)[names(OS_Ht) == "value"] <- "Mean_Height"
OS_Ht <- subset(OS_Ht, select = -c(variable))

OS_CBH <- OS_Ht_CBH_melt[grep(pattern = "CBH", OS_Ht_CBH_melt$variable, fixed = TRUE),]
names(OS_CBH)[names(OS_CBH) == "value"] <- "Mean_CBH"
OS_CBH <- subset(OS_CBH, select = -c(variable))

OS_output <- merge(OS_CBH, OS_Ht, all = TRUE)

# Species per hectare

OS_SPH_melt <- reshape2::melt(OS_SPH, id = "DBH.Class")
setnames(OS_SPH_melt, old = c("variable", "value"), new = c("Species", "SPH"))

OS_output <- merge(OS_output, OS_SPH_melt,  by =  c("Species", "DBH.Class"), all = TRUE)

# Basal area

OS_BA_melt <- reshape2::melt(OS_BA, id = "DBH.Class")
setnames(OS_BA_melt, old = c("variable", "value"), new = c("Species", "BA"))

OS_output <- merge(OS_output, OS_BA_melt,  by =  c("Species", "DBH.Class"), all = TRUE)

# Volume

OS_Vol_melt <- reshape2::melt(OS_Vol, id = "DBH.Class")
setnames(OS_Vol_melt, old = c("variable", "value"), new = c("Species", "Volume"))

OS_output <- merge(OS_output, OS_Vol_melt,  by =  c("Species", "DBH.Class"), all = TRUE)

###########

setnames(OS_output, old = c("DBH.Class"), new = c("DBH_class"))

OS_output[OS_output == "12.5 - 17.5"] = "Layer 1 (12.5 - 17.5 cm dbh)"
OS_output[OS_output == "17.5 - 22.5"] = "Layer 1 (17.5 - 22.5 cm dbh)"
OS_output[OS_output == "22.5 - 27.5"] = "Layer 1 (22.5 - 27.5 cm dbh)"
OS_output[OS_output == "27.5 - 35"] = "Layer 1 (27.5 - 35 cm dbh)"
OS_output[OS_output == "35 - 45"] = "Layer 1 (35 - 45 cm dbh)"
OS_output[OS_output == "45+"] = "Layer 1 (45+ cm dbh)"

# 2. Understorey

# Height and CBH

US_Ht_CBH_melt <- reshape2::melt(US_Ht_CBH, id = "Layer")
US_Ht_CBH_melt$Species <- gsub("\\.", "", substr(US_Ht_CBH_melt$variable, start = 1, stop = 3))


#Check for yew
US_Ht_CBH_melt$Species[US_Ht_CBH_melt$Species == "T."] <- "T"
US_Ht_CBH_melt$Species[US_Ht_CBH_melt$Species == "Dea"] <- "DP"

US_Ht <- US_Ht_CBH_melt[grep(pattern = "Ht", US_Ht_CBH_melt$variable, fixed = TRUE),]
names(US_Ht)[names(US_Ht) == "value"] <- "Mean_Height"
US_Ht <- subset(US_Ht, select = -c(variable))

US_CBH <- US_Ht_CBH_melt[grep(pattern = "CBH", US_Ht_CBH_melt$variable, fixed = TRUE),]
names(US_CBH)[names(US_CBH) == "value"] <- "Mean_CBH"
US_CBH <- subset(US_CBH, select = -c(variable))

US_output <- merge(US_CBH, US_Ht, all = TRUE)

# Species per hectare
  #change Dead to DP
if ("Dead" %in% colnames(US_SPH)) {
  colnames(US_SPH)[colnames(US_SPH) == "Dead"] <- "DP"
}

US_SPH_melt <- reshape2::melt(US_SPH, id = "Layer")
setnames(US_SPH_melt, old = c("variable", "value"), new = c("Species", "SPH"))

US_output <- merge(US_output, US_SPH_melt,  by =  c("Species", "Layer"), all = TRUE)

##############

US_output$BA <- NA
US_output$Volume <- NA

setnames(US_output, old = c("Layer"), new = c("DBH_class"))
US_output[US_output == "Layer 1 (12.5-17.5)"] = "Layer 1 (12.5 - 17.5 cm dbh)"

# Joining overstorey and understorey

All_output <- rbind(OS_output, US_output)
All_output <- All_output[!duplicated(All_output),]

  # Will depend whether Dead Useless or Dead Potential 
All_output[All_output == "Dead"] = "DU"

  # Remove totals for species across all diameter classes
All_output <- All_output[which(All_output$DBH_class != "Total"),]
  #remove rows where sph = 0
All_output <- All_output %>% filter(Species != "X")
All_output <- All_output %>% filter(Species !="Total")

# Joining the cutting specs

cuttingSpecs <- subset(cuttingSpecs, select = -c(Stand.Layer))
cuttingSpecs_melt <- reshape2::melt(cuttingSpecs, id = "DBH.Class")
setnames(cuttingSpecs_melt, old = c("DBH.Class", "variable", "value"), new = c("DBH_class", "Species", "Cutting_Spec"))

cuttingSpecs_melt[cuttingSpecs_melt == "0 - 1.5"] = "Layer 4 (<1.3m)"
cuttingSpecs_melt[cuttingSpecs_melt == "1.5 - 7.5"] = "Layer 3 (2.5-7.49)"
cuttingSpecs_melt[cuttingSpecs_melt == "7.5 - 12.5"] = "Layer 2 (7.5-12.49)"
cuttingSpecs_melt[cuttingSpecs_melt == "12.5 - 17.5"] = "Layer 1 (12.5 - 17.5 cm dbh)"
cuttingSpecs_melt[cuttingSpecs_melt == "17.5 - 22.5"] = "Layer 1 (17.5 - 22.5 cm dbh)"
cuttingSpecs_melt[cuttingSpecs_melt == "22.5 - 27.5"] = "Layer 1 (22.5 - 27.5 cm dbh)"
cuttingSpecs_melt[cuttingSpecs_melt == "27.5 - 35"] = "Layer 1 (27.5 - 35 cm dbh)"
cuttingSpecs_melt[cuttingSpecs_melt == "35 - 45"] = "Layer 1 (35 - 45 cm dbh)"
cuttingSpecs_melt[cuttingSpecs_melt == "45+"] = "Layer 1 (45+ cm dbh)"

All_output <- merge(All_output, cuttingSpecs_melt,  by =  c("Species", "DBH_class"), all.x = TRUE)

# Calculating cut and leave
All_output <- All_output %>% mutate(SPH_cut = round(SPH*Cutting_Spec/100, digits = 0)) %>%
                             mutate(SPH_leave = round(SPH*(100-Cutting_Spec)/100, digits = 0)) %>%
                             mutate(Volume_cut = round(Volume*Cutting_Spec/100, digits = 0)) %>%
                             mutate(Volume_leave = round(Volume*(100-Cutting_Spec)/100, digits = 0)) %>%
                             mutate(BA_cut = round(BA*Cutting_Spec/100, digits = 0)) %>%
                             mutate(BA_leave = round(BA*(100-Cutting_Spec)/100, digits = 0))


All_output <- All_output[!is.na(All_output$SPH),]
All_output<-All_output[All_output$SPH != 0,]
All_output <- All_output[!All_output$Species == "Total",]
All_output[is.na(All_output)]<-0

# Set height values for dead trees

All_output[which(All_output$Species == "DP" & All_output$DBH_class == "Layer 1 (12.5 - 17.5 cm dbh)"), c("Mean_Height")] <- 8
All_output[which(All_output$Species == "DU" & All_output$DBH_class == "Layer 3 (2.5-7.49)"), c("Mean_Height")] <- 2
All_output[which(All_output$Species == "DU" & All_output$DBH_class == "Layer 4 (<1.3m)"), c("Mean_Height")] <- 1

# Calculating Totals

All_output <- mutate(All_output, Decay = ifelse(All_output$Species == "DP" | All_output$Species == "DU", "Dead", "Live"))
All_output <- mutate(All_output, Conifer = ifelse(All_output$Species %in% coniferList, "Coniferous", "Other"))

All_output$Mean_CBH[All_output$Conifer != "Coniferous"] <- NA

#
All_output$Mean_CBH<-as.numeric(All_output$Mean_CBH)
All_output$Mean_Height<-as.numeric(All_output$Mean_Height)
#

All_output <- mutate(All_output, CBH_forWtMean = All_output$Mean_CBH*All_output$SPH)

All_output <- mutate(All_output, Ht_forWtMean = All_output$Mean_Height*All_output$SPH)
All_output$CBH_forWtMean[is.na(All_output$CBH_forWtMean)]<-0
All_output$Ht_forWtMean[is.na(All_output$Ht_forWtMean)]<-0
# Conifer

coniferTemp <- aggregate(cbind(SPH, SPH_cut, SPH_leave,
                Volume, Volume_cut, Volume_leave,
                BA, BA_cut, BA_leave, CBH_forWtMean, Ht_forWtMean) ~ Conifer+DBH_class,
                data = All_output, FUN = sum, na.action = na.pass)

coniferTemp <- coniferTemp %>% 
  filter(Conifer == "Coniferous") %>%
  mutate(Species = "Total Conifer") %>%
  mutate(Mean_CBH = round(CBH_forWtMean/SPH, digits = 1)) %>%
  mutate(Mean_Height = round(Ht_forWtMean/SPH, digits = 1)) %>%
  dplyr::select(-Conifer, -CBH_forWtMean, -Ht_forWtMean)

existingLayers <- c("Layer 4 (<1.3m)",
                    "Layer 3 (2.5-7.49)",
                    "Layer 2 (7.5-12.49)",
                    "Layer 1 (12.5 - 17.5 cm dbh)",
                    "Layer 1 (17.5 - 22.5 cm dbh)",
                    "Layer 1 (22.5 - 27.5 cm dbh)",
                    "Layer 1 (27.5 - 35 cm dbh)",
                    "Layer 1 (35 - 45 cm dbh)",
                    "Layer 1 (45+ cm dbh)")

coniferLayers <- unique(coniferTemp$DBH_class)
missingLayers <- existingLayers[!(existingLayers %in% coniferLayers)]

if(length(missingLayers) > 0){

  coniferTempFill <- data.frame(Species = "Total Conifer", DBH_class = missingLayers,
                       SPH = 0, SPH_cut = 0, SPH_leave = 0,
                       Volume = 0, Volume_cut = 0, Volume_leave = 0,
                       BA = 0, BA_cut = 0, BA_leave = 0,
                       Mean_CBH = NA, Mean_Height = NA)
  
  coniferTemp <- rbind.fill(coniferTemp, coniferTempFill)
}
  
# All Live

liveTemp <- aggregate(cbind(SPH, SPH_cut, SPH_leave,
                               Volume, Volume_cut, Volume_leave,
                               BA, BA_cut, BA_leave, CBH_forWtMean, Ht_forWtMean) ~ Decay+DBH_class,
                               data = All_output, FUN = sum, na.action = na.pass)

liveTemp <- liveTemp %>% 
  filter(Decay == "Live") %>%
  mutate(Species = "Total Live") %>%
  mutate(Mean_CBH = round(CBH_forWtMean/SPH, digits = 1)) %>%
  mutate(Mean_Height = round(Ht_forWtMean/SPH, digits = 1)) %>%
  dplyr::select(-Decay, -CBH_forWtMean, -Ht_forWtMean)


liveLayers <- unique(liveTemp$DBH_class)
missingLayers <- existingLayers[!(existingLayers %in% liveLayers)]

if(length(missingLayers) > 0){
  
  liveTempFill <- data.frame(Species = "Total Live", DBH_class = missingLayers,
                                SPH = 0, SPH_cut = 0, SPH_leave = 0,
                                Volume = 0, Volume_cut = 0, Volume_leave = 0,
                                BA = 0, BA_cut = 0, BA_leave = 0,
                                Mean_CBH = NA, Mean_Height = NA)
  
  liveTemp <- rbind.fill(liveTemp, liveTempFill)
}

  # Everything per layer

allPerLayerTemp <- aggregate(cbind(SPH, SPH_cut, SPH_leave,
                               Volume, Volume_cut, Volume_leave,
                               BA, BA_cut, BA_leave, CBH_forWtMean, Ht_forWtMean) ~ DBH_class,
                         data = All_output, FUN = sum, na.action = na.pass)

allPerLayerTemp <- allPerLayerTemp %>% mutate(Species = "Total All Species") %>%
            mutate(Mean_CBH = round(CBH_forWtMean/SPH, digits = 1)) %>%
            mutate(Mean_Height = round(Ht_forWtMean/SPH, digits = 1)) %>%
  dplyr::select(-CBH_forWtMean, -Ht_forWtMean)

allLayers <- unique(allPerLayerTemp$DBH_class)
missingLayers <- existingLayers[!(existingLayers %in% allLayers)]

if(length(missingLayers) > 0){
  
  allPerLayerTempFill <- data.frame(Species = "Total All Species", DBH_class = missingLayers,
                             SPH = 0, SPH_cut = 0, SPH_leave = 0,
                             Volume = 0, Volume_cut = 0, Volume_leave = 0,
                             BA = 0, BA_cut = 0, BA_leave = 0,
                             Mean_CBH = NA, Mean_Height = NA)
  
  allPerLayerTemp <- rbind.fill(allPerLayerTemp, allPerLayerTempFill)
}


  # Everything per layer

allTemp <- data.frame(Species = "Total All Layers", DBH_class = NA,
                                  SPH = sum(All_output$SPH, na.rm = T), SPH_cut = sum(All_output$SPH_cut, na.rm = T), SPH_leave = sum(All_output$SPH_leave, na.rm = T),
                                  Volume = sum(All_output$Volume, na.rm = T), Volume_cut = sum(All_output$Volume_cut, na.rm = T), Volume_leave = sum(All_output$Volume_leave, na.rm = T),
                                  BA = sum(All_output$BA, na.rm = T), BA_cut = sum(All_output$BA_cut, na.rm = T), BA_leave = sum(All_output$BA_leave, na.rm = T),
                                  CBH_forWtMean = sum(All_output$CBH_forWtMean), Ht_forWtMean = sum(All_output$Ht_forWtMean, na.rm = T))

allTemp <- allTemp %>% mutate(Mean_CBH=round(CBH_forWtMean/SPH,digits = 1),
                              Mean_Height = round(Ht_forWtMean/SPH, digits = 1)) %>%
  dplyr::select(-c("CBH_forWtMean","Ht_forWtMean"))
  # Layer-1 specific

L1temp <- All_output[grep(pattern = "Layer 1", All_output$DBH_class, fixed = TRUE),]

    # Conifers

L1conifer <- aggregate(cbind(SPH, SPH_cut, SPH_leave,
                           Volume, Volume_cut, Volume_leave,
                           BA, BA_cut, BA_leave, CBH_forWtMean, Ht_forWtMean) ~ Conifer,
                     data = L1temp, FUN = sum, na.rm = TRUE, na.action = na.pass)

L1conifer <- L1conifer %>% filter(Conifer == "Coniferous") %>%
              mutate(Species = "Total Layer 1 - Conifers Only") %>%
              mutate(DBH_class = "Total Layer 1") %>%
              mutate(Mean_CBH = round(CBH_forWtMean/SPH, digits = 1)) %>%
              mutate(Mean_Height = round(Ht_forWtMean/SPH, digits = 1)) %>%
  dplyr::select(-c("Conifer", "CBH_forWtMean", "Ht_forWtMean"))

    # All Species

L1all <- L1temp %>% dplyr::select(SPH, SPH_cut, SPH_leave,
                           Volume, Volume_cut, Volume_leave,
                           BA, BA_cut, BA_leave, CBH_forWtMean, Ht_forWtMean)

L1all <- as.data.frame(t(colSums(L1all, na.rm = TRUE)))

L1all <- L1all %>% mutate(Species = "Total Layer 1 - All Species") %>%
                    mutate(DBH_class = "Total Layer 1") %>%
                    mutate(Mean_CBH = round(CBH_forWtMean/SPH, digits = 1)) %>%
                    mutate(Mean_Height = round(Ht_forWtMean/SPH, digits = 1)) %>%
  dplyr::select(-c("CBH_forWtMean", "Ht_forWtMean"))

    # Dead Potential

existingDPlayers <- All_output$DBH_class[All_output$Species == "DP"]

allDPlayers <- c("Layer 2 (7.5-12.49)",
                 "Layer 1 (12.5 - 17.5 cm dbh)",
                 "Layer 1 (17.5 - 22.5 cm dbh)",
                 "Layer 1 (22.5 - 27.5 cm dbh)",
                 "Layer 1 (27.5 - 35 cm dbh)",
                 "Layer 1 (35 - 45 cm dbh)",
                 "Layer 1 (45+ cm dbh)")

missingDPlayers <- allDPlayers[!(allDPlayers %in% existingDPlayers)]

DPtemp <- data.frame(Species = "DP", DBH_class = missingDPlayers,
           SPH = 0, SPH_cut = 0, SPH_leave = 0,
           Volume = 0, Volume_cut = 0, Volume_leave = 0,
           BA = 0, BA_cut = 0, BA_leave = 0)

# Put everything together...

All_output <- rbind.fill(All_output, coniferTemp, liveTemp, allPerLayerTemp, allTemp, L1conifer, L1all, DPtemp)

# Reformatting

All_output <- All_output[,c("Species", "DBH_class", "Mean_CBH", "Mean_Height",
                            "SPH", "SPH_cut", "SPH_leave",
                            "Volume", "Volume_cut", "Volume_leave",
                            "BA", "BA_cut", "BA_leave",
                            "Cutting_Spec")]

setnames(All_output, old = c("SPH", "Volume", "BA", "Cutting_Spec"), new = c("SPH_exist", "Volume_exist", "BA_exist", "Cutting_spec"))

All_output$Species[All_output$Species == "DP"] <- "Total Dead Potential"
All_output$DBH_class[All_output$DBH_class == "Layer 2 (7.5-12.49)"] <- "Layer 2 (7.5 - 12.49 cm dbh)"
All_output$DBH_class[All_output$DBH_class == "Layer 3 (2.5-7.49)"] <- "Layer 3 (2.5 - 7.49 cm dbh)"
All_output$DBH_class[All_output$DBH_class == "Layer 4 (<1.3m)"] <- "Layer 4 (<1.3 cm dbh)"

All_output$Cutting_spec[is.na(All_output$Cutting_spec)] <- 0

#changed to keep cbh calculation for totals despite only including conifers
totallingLayers <- c("Total All Species",
                     "Total Dead Potential",
                     "Total Live",
                 "Total Layer 1 - All Species")

#totallingLayers <- c( "Total Dead Potential","Total Layer 1 - All Species")

VolBA <- c("Volume_exist",
           "Volume_cut",
           "Volume_leave",
           "BA_exist",
           "BA_cut",
           "BA_leave")

USlayers <- c("Layer 2 (7.5 - 12.49 cm dbh)",
              "Layer 3 (2.5 - 7.49 cm dbh)",
              "Layer 4 (<1.3 cm dbh)")

#All_output$Mean_CBH[which(All_output$Species %in% totallingLayers)] <- "-"

All_output[which(All_output$DBH_class %in% USlayers), which(colnames(All_output) %in% VolBA)] <- "-"

#All_output[is.na(All_output)] <- "-"

All_output$Species <- factor(All_output$Species, levels = c(coniferList,
                                                            nonConiferList[!nonConiferList == "DP"],
                                                            "Total Dead Potential",
                                                            "Total Live",
                                                            "Total All Species",
                                                            "Total Layer 1 - All Species",
                                                            "Total Conifer",
                                                            "Total Layer 1 - Conifers Only"), ordered = TRUE)

All_output$DBH_class <- factor(All_output$DBH_class, levels = c("Layer 1 (45+ cm dbh)",
                                                                "Layer 1 (35 - 45 cm dbh)",
                                                                "Layer 1 (27.5 - 35 cm dbh)",
                                                                "Layer 1 (22.5 - 27.5 cm dbh)",
                                                                "Layer 1 (17.5 - 22.5 cm dbh)",
                                                                "Layer 1 (12.5 - 17.5 cm dbh)",
                                                                "Total Layer 1",
                                                                "Layer 2 (7.5 - 12.49 cm dbh)",
                                                                "Layer 3 (2.5 - 7.49 cm dbh)",
                                                                "Layer 4 (<1.3 cm dbh)"), ordered = TRUE)

All_output <- All_output[with(All_output, order(DBH_class, Species)),]
All_output$Species <- as.character(All_output$Species)
All_output$DBH_class <- as.character(All_output$DBH_class)

# Extract layer totals

allTemp <- tail(All_output, n = 1)
All_output <- slice(All_output, 1:(n() - 1)) 

allTemp$Species <- "TOTAL ALL LAYERS"
allTemp$DBH_class <- "Layer Totals"


USlayers <- c("Layer 2 (7.5 - 12.49 cm dbh)",
              "Layer 3 (2.5 - 7.49 cm dbh)",
              "Layer 4 (<1.3 cm dbh)")

totalsL1 <- All_output[All_output$Species == "Total Layer 1 - All Species" & All_output$DBH_class == "Total Layer 1",]
totalsUS <- All_output[All_output$Species == "Total All Species" & All_output$DBH_class %in% USlayers,]

totals <- rbind.fill(totalsL1,totalsUS,allTemp)
totals <- totals[grepl("^\\d+\\.?\\d*$", totals$SPH_exist), ]
totals
# Format and append totals

totals$Species <- c("Total All Species Layer 1",
                    "Total All Species Layer 2",
                    "Total All Species Layer 3",
                    "Total All Species Layer 4",
                    "TOTAL ALL LAYERS")

totals$DBH_class <- "Layer Totals"

All_output <- rbind.fill(All_output,totals)
All_output[is.na(All_output)] <- "-"

#The code at this point will automatically write out the fused table as a csv, and example can be found here:
# Z:\Admin\References\Methods\SNAP_to_BCWS_tablesexample_bsws_table_unformatted.csv
write.csv(All_output,file = paste0(path,s_s_prefix,table_folder,TR,"_SST.csv"),row.names = F)

#####################################################################################################################




{
  data.stand <- read.csv(paste0(path,s_s_prefix,table_folder,TR,"_SST.csv"), header = TRUE) #specify directory where you saved csv file
  #update start & end row number for the output_table code block (starting at line 19) by viewing the data.stand, 
  ###then run the following code###
  data.stand.final <- dplyr::select(data.stand,-c(DBH_class,Cutting_spec))
  colnames(data.stand.final) <- c("Species","Average Crown to Base Height (m)",
                                  "Average Tree Height (m)", "Existing", "Cut",
                                  "Leave", "Existing", "Cut",
                                  "Leave","Existing", "Cut",
                                  "Leave")
  
  #csv, not neccessary
  #write.csv(data.stand.final, file = paste0(path,"/Outputs/",TR,"_SST.csv"), row.names = TRUE)
  ########################start formatting the output table###########################################################################
  #please note that the start and end row number for each layer need to be updated before running the code below
  output_table <- kableExtra::kbl(data.stand.final, caption = "", align = "c")%>%
    kableExtra::kable_paper(full_width = F)%>%
    kableExtra::pack_rows("Layer 1 (>45 cm dbh)", 
              which(data.stand$DBH_class=="Layer 1 (45+ cm dbh)")[1],
              (which(data.stand$DBH_class=="Layer 1 (45+ cm dbh)")[(length(which(data.stand$DBH_class=="Layer 1 (45+ cm dbh)")))]),
              label_row_css = "background-color: #777; color: #fff;")%>%
    kableExtra::pack_rows("Layer 1 (35 - 45 cm dbh)",
              which(data.stand$DBH_class=="Layer 1 (35 - 45 cm dbh)")[1],
              (which(data.stand$DBH_class=="Layer 1 (35 - 45 cm dbh)")[(length(which(data.stand$DBH_class=="Layer 1 (35 - 45 cm dbh)")))]),
              label_row_css = "background-color: #777; color: #fff;")%>%
    kableExtra::pack_rows("Layer 1 (27.5 - 35 cm dbh)",
              which(data.stand$DBH_class=="Layer 1 (27.5 - 35 cm dbh)")[1],
              (which(data.stand$DBH_class=="Layer 1 (27.5 - 35 cm dbh)")[(length(which(data.stand$DBH_class=="Layer 1 (27.5 - 35 cm dbh)")))]),
              label_row_css = "background-color: #777; color: #fff;")%>%
    kableExtra::pack_rows("Layer 1 (22.5 - 27.5 cm dbh)", 
              which(data.stand$DBH_class=="Layer 1 (22.5 - 27.5 cm dbh)")[1],
              (which(data.stand$DBH_class=="Layer 1 (22.5 - 27.5 cm dbh)")[(length(which(data.stand$DBH_class=="Layer 1 (22.5 - 27.5 cm dbh)")))]),
              label_row_css = "background-color: #777; color: #fff;")%>%
    kableExtra::pack_rows("Layer 1 (17.5 - 22.5 cm dbh)",
              which(data.stand$DBH_class=="Layer 1 (17.5 - 22.5 cm dbh)")[1],
              (which(data.stand$DBH_class=="Layer 1 (17.5 - 22.5 cm dbh)")[(length(which(data.stand$DBH_class=="Layer 1 (17.5 - 22.5 cm dbh)")))]),
              label_row_css = "background-color: #777; color: #fff;")%>%
    kableExtra::pack_rows("Layer 1 (12.5 - 17.5 cm dbh)", 
              which(data.stand$DBH_class=="Layer 1 (12.5 - 17.5 cm dbh)")[1],
              (which(data.stand$DBH_class=="Layer 1 (12.5 - 17.5 cm dbh)")[(length(which(data.stand$DBH_class=="Layer 1 (12.5 - 17.5 cm dbh)")))]),
              label_row_css = "background-color: #777; color: #fff;")%>%
    kableExtra::pack_rows("Total Layer 1", 
              which(data.stand$DBH_class=="Total Layer 1")[1],
              (which(data.stand$DBH_class=="Total Layer 1")[(length(which(data.stand$DBH_class=="Total Layer 1")))]),
              label_row_css = "background-color: #777; color: #fff;")%>%
    kableExtra::pack_rows("Layer 2 (7.5 - 12.49 cm dbh)", 
              which(data.stand$DBH_class=="Layer 2 (7.5 - 12.49 cm dbh)")[1],
              (which(data.stand$DBH_class=="Layer 2 (7.5 - 12.49 cm dbh)")[(length(which(data.stand$DBH_class=="Layer 2 (7.5 - 12.49 cm dbh)")))]),
              label_row_css = "background-color: #777; color: #fff;")%>%
    kableExtra::pack_rows("Layer 3 (2.5 - 7.49 cm dbh)",
              which(data.stand$DBH_class=="Layer 3 (2.5 - 7.49 cm dbh)")[1],
              (which(data.stand$DBH_class=="Layer 3 (2.5 - 7.49 cm dbh)")[(length(which(data.stand$DBH_class=="Layer 3 (2.5 - 7.49 cm dbh)")))]),
              label_row_css = "background-color: #777; color: #fff;")%>%
    kableExtra::pack_rows("Layer 4 (<1.3 cm dbh)",
              which(data.stand$DBH_class=="Layer 4 (<1.3 cm dbh)")[1],
              (which(data.stand$DBH_class=="Layer 4 (<1.3 cm dbh)")[(length(which(data.stand$DBH_class=="Layer 4 (<1.3 cm dbh)")))]),
              label_row_css = "background-color: #777; color: #fff;")%>%
    kableExtra::pack_rows("Layer Totals",
              which(data.stand$DBH_class=="Layer Totals")[1],
              (which(data.stand$DBH_class=="Layer Totals")[(length(which(data.stand$DBH_class=="Layer Totals")))]),
              label_row_css = "background-color: #777; color: #fff;") %>%
    kableExtra::add_header_above(c("","","", "STEM PER HECTARE (SPH)"=3, "VOLUME (m3/ha)"=3, "BASAL AREA (m2/ha)"=3))%>%
    kableExtra::add_header_above(c("STAND AND STOCK TABLE "=12),
                     align = "left")
 
  #html
  kableExtra::save_kable(output_table,file=paste0(path,s_s_prefix,table_folder,TR,"_SST.html"))

 
}
 #Export to Excel
# Initialize a list to store the formatted table
formatted_table <- list()

# Layer 1 (>45 cm dbh)
formatted_table[[1]] <- data.frame(
  Species = "Layer 1 (>45 cm dbh)", `Average Crown to Base Height (m)` = NA, `Average Tree Height (m)` = NA,
  Existing = NA, Cut = NA, Leave = NA, Existing = NA, Cut = NA, Leave = NA, Existing = NA, Cut = NA, Leave = NA,check.names = FALSE
)
formatted_table[[2]] <- data.stand.final[which(data.stand$DBH_class == "Layer 1 (45+ cm dbh)"), ]

# Layer 1 (35 - 45 cm dbh)
formatted_table[[3]] <- data.frame(
  Species = "Layer 1 (35 - 45 cm dbh)", `Average Crown to Base Height (m)` = NA, `Average Tree Height (m)` = NA,
 Existing = NA, Cut = NA, Leave = NA, Existing = NA, Cut = NA, Leave = NA, Existing = NA, Cut = NA, Leave = NA,check.names = FALSE
)
formatted_table[[4]] <- data.stand.final[which(data.stand$DBH_class == "Layer 1 (35 - 45 cm dbh)"), ]

# Layer 1 (27.5 - 35 cm dbh)
formatted_table[[5]] <- data.frame(
  Species = "Layer 1 (27.5 - 35 cm dbh)", `Average Crown to Base Height (m)` = NA, `Average Tree Height (m)` = NA,
  Existing = NA, Cut = NA, Leave = NA, Existing = NA, Cut = NA, Leave = NA, Existing = NA, Cut = NA, Leave = NA,check.names = FALSE
)
formatted_table[[6]] <- data.stand.final[which(data.stand$DBH_class == "Layer 1 (27.5 - 35 cm dbh)"), ]

# Layer 1 (22.5 - 27.5 cm dbh)
formatted_table[[7]] <- data.frame(
  Species = "Layer 1 (22.5 - 27.5 cm dbh)", `Average Crown to Base Height (m)` = NA, `Average Tree Height (m)` = NA,
  Existing = NA, Cut = NA, Leave = NA, Existing = NA, Cut = NA, Leave = NA, Existing = NA, Cut = NA, Leave = NA,check.names = FALSE
)
formatted_table[[8]] <- data.stand.final[which(data.stand$DBH_class == "Layer 1 (22.5 - 27.5 cm dbh)"), ]

# Layer 1 (17.5 - 22.5 cm dbh)
formatted_table[[9]] <- data.frame(
  Species = "Layer 1 (17.5 - 22.5 cm dbh)", `Average Crown to Base Height (m)` = NA, `Average Tree Height (m)` = NA,
  Existing = NA, Cut = NA, Leave = NA, Existing = NA, Cut = NA, Leave = NA, Existing = NA, Cut = NA, Leave = NA,check.names = FALSE
)
formatted_table[[10]] <- data.stand.final[which(data.stand$DBH_class == "Layer 1 (17.5 - 22.5 cm dbh)"), ]

# Layer 1 (12.5 - 17.5 cm dbh)
formatted_table[[11]] <- data.frame(
  Species = "Layer 1 (12.5 - 17.5 cm dbh)", `Average Crown to Base Height (m)` = NA, `Average Tree Height (m)` = NA,
  Existing = NA, Cut = NA, Leave = NA, Existing = NA, Cut = NA, Leave = NA, Existing = NA, Cut = NA, Leave = NA,check.names = FALSE
)
formatted_table[[12]] <- data.stand.final[which(data.stand$DBH_class == "Layer 1 (12.5 - 17.5 cm dbh)"), ]

# Total Layer 1
formatted_table[[13]] <- data.frame(
  Species = "Total Layer 1", `Average Crown to Base Height (m)` = NA, `Average Tree Height (m)` = NA,
  Existing = NA, Cut = NA, Leave = NA, Existing = NA, Cut = NA, Leave = NA, Existing = NA, Cut = NA, Leave = NA,check.names = FALSE
)
formatted_table[[14]] <- data.stand.final[which(data.stand$DBH_class == "Total Layer 1"), ]

# Layer 2 (7.5 - 12.49 cm dbh)
formatted_table[[15]] <- data.frame(
  Species = "Layer 2 (7.5 - 12.49 cm dbh)", `Average Crown to Base Height (m)` = NA, `Average Tree Height (m)` = NA,
  Existing = NA, Cut = NA, Leave = NA, Existing = NA, Cut = NA, Leave = NA, Existing = NA, Cut = NA, Leave = NA,check.names = FALSE
)
formatted_table[[16]] <- data.stand.final[which(data.stand$DBH_class == "Layer 2 (7.5 - 12.49 cm dbh)"), ]

# Layer 3 (2.5 - 7.49 cm dbh)
formatted_table[[17]] <- data.frame(
  Species = "Layer 3 (2.5 - 7.49 cm dbh)", `Average Crown to Base Height (m)` = NA, `Average Tree Height (m)` = NA,
  Existing = NA, Cut = NA, Leave = NA, Existing = NA, Cut = NA, Leave = NA, Existing = NA, Cut = NA, Leave = NA,check.names = FALSE
)
formatted_table[[18]] <- data.stand.final[which(data.stand$DBH_class == "Layer 3 (2.5 - 7.49 cm dbh)"), ]

# Layer 4 (<1.3 cm dbh)
formatted_table[[19]] <- data.frame(
  Species = "Layer 4 (<1.3 cm dbh)", `Average Crown to Base Height (m)` = NA, `Average Tree Height (m)` = NA,
  Existing = NA, Cut = NA, Leave = NA, Existing = NA, Cut = NA, Leave = NA, Existing = NA, Cut = NA, Leave = NA,check.names = FALSE
)
formatted_table[[20]] <- data.stand.final[which(data.stand$DBH_class == "Layer 4 (<1.3 cm dbh)"), ]

# Layer Totals
formatted_table[[21]] <- data.frame(
  Species = "Layer Totals", `Average Crown to Base Height (m)` = NA, `Average Tree Height (m)` = NA,
  Existing = NA, Cut = NA, Leave = NA, Existing = NA, Cut = NA, Leave = NA, Existing = NA, Cut = NA, Leave = NA,check.names = FALSE
)
formatted_table[[22]] <- data.stand.final[which(data.stand$DBH_class == "Layer Totals"), ]

# Combine all parts into a single data frame
final_output <- do.call(rbind, formatted_table)

# Create a new Excel workbook
wb <- openxlsx::createWorkbook()

# Add a worksheet to the workbook
openxlsx::addWorksheet(wb, "Stand and Stock Table")

# Write the final data frame to the worksheet
  openxlsx::writeData(wb, "Stand and Stock Table", final_output,startRow = 2)
library(openxlsx)
# Merge cells and add headers above the microheaders
  mergeCells(wb, "Stand and Stock Table", cols = 1:3, rows = 1)
  mergeCells(wb, "Stand and Stock Table", cols = 4:6, rows = 1)
  mergeCells(wb, "Stand and Stock Table", cols = 7:9, rows = 1)
  mergeCells(wb, "Stand and Stock Table", cols = 10:12, rows = 1)

# Add the text for the headers
  writeData(wb, "Stand and Stock Table", "Stand and Stock Table", startCol = 1, startRow = 1)
  writeData(wb, "Stand and Stock Table", "STEM PER HECTARE (SPH)", startCol = 4, startRow = 1)
  writeData(wb, "Stand and Stock Table", "VOLUME (m3/ha)", startCol = 7, startRow = 1)
  writeData(wb, "Stand and Stock Table", "BASAL AREA (m2/ha)", startCol = 10, startRow = 1)

# Format the headers (bold, size 14)
  headerStyle <- createStyle(fontSize = 14, textDecoration = "bold", halign = "center")
  addStyle(wb, "Stand and Stock Table", headerStyle, rows = 1, cols = 1:3, gridExpand = TRUE)
  addStyle(wb, "Stand and Stock Table", headerStyle, rows = 1, cols = 4:6, gridExpand = TRUE)
  addStyle(wb, "Stand and Stock Table", headerStyle, rows = 1, cols = 7:9, gridExpand = TRUE)
  addStyle(wb, "Stand and Stock Table", headerStyle, rows = 1, cols = 10:12, gridExpand = TRUE)

# Add bold style for the subheaders (row 2)
  subheaderStyle <- createStyle(textDecoration = "bold", halign = "center",wrapText = TRUE, valign = "center")
  addStyle(wb, "Stand and Stock Table", subheaderStyle, rows = 2, cols = 1:ncol(final_output), gridExpand = TRUE)

# Center-align all numeric columns (excluding the first column)
numericColumnStyle <- createStyle(halign = "center")
  # Apply center alignment to all rows starting from row 3 (where data begins), from column 2 to the last column
  addStyle(wb, "Stand and Stock Table", numericColumnStyle, rows = 3:(nrow(final_output) + 2), cols = 2:ncol(final_output), gridExpand = TRUE)
# Define the rows that need to be bolded in the Species column
bold_rows <- c("Layer 1 (>45 cm dbh)", "Layer 1 (35 - 45 cm dbh)", "Layer 1 (27.5 - 35 cm dbh)",
               "Layer 1 (22.5 - 27.5 cm dbh)", "Layer 1 (17.5 - 22.5 cm dbh)", "Layer 1 (12.5 - 17.5 cm dbh)",
               "Total Layer 1", "Layer 2 (7.5 - 12.49 cm dbh)", "Layer 3 (2.5 - 7.49 cm dbh)", 
               "Layer 4 (<1.3 cm dbh)", "Layer Totals", "TOTAL ALL LAYERS")

# Create a bold style for the specific rows in the Species column
boldStyle <- createStyle(textDecoration = "bold")

# Loop through the final_output data frame and apply the bold style to the matching rows in the Species column
for (x in 1:nrow(final_output)) {
  if (final_output$Species[x] %in% bold_rows) {
    # Apply the bold style to the Species column for the matching rows (column 1)
    addStyle(wb, "Stand and Stock Table", boldStyle, rows = x + 2, cols = 1, gridExpand = TRUE)
  }
}
# Save the workbook to an Excel file
saveWorkbook(wb, paste0(path,"/Outputs/",TR,"_SST.xlsx"), overwrite = TRUE)

}

####################################################End##################################################################################
```


  6.2 Run Code below if you have NO OVERSTORY: But check results thoroughly: either adjust code to loop over treatments or run it specifically for the treatment unit number without overstory

```{r}
# Number of unique stratum
stratas <- unique(Snap_EX$Stratum)

# Iterate over treatments
#for(i in 1:stratas){

#Set Strata you are working on:-------------------------------
i<-2
#:----------------------------------------------------------------

  TR <- stratas[i]
  table_folder <- paste0(TR, "_tables/")
  
  # Read CSV files
  cuttingSpecs <- read.csv(file = paste0(path, s_s_prefix, table_folder, "cuttingSpecs.csv"), row.names = NULL, stringsAsFactors = FALSE)
  OS_SPH <- read.csv(file = paste0(path, s_s_prefix, table_folder, "OS_SPH.csv"), row.names = NULL, stringsAsFactors = FALSE)
  OS_BA <- read.csv(file = paste0(path, s_s_prefix, table_folder, "OS_BA.csv"), row.names = NULL, stringsAsFactors = FALSE)
  OS_Vol <- read.csv(file = paste0(path, s_s_prefix, table_folder, "OS_Vol.csv"), row.names = NULL, stringsAsFactors = FALSE)
  OS_Ht_CBH <- read.csv(file = paste0(path, s_s_prefix, table_folder, "OS_Ht_CBH.csv"), row.names = NULL, stringsAsFactors = FALSE)
  US_SPH <- read.csv(file = paste0(path, s_s_prefix, table_folder, "US_SPH.csv"), row.names = NULL, stringsAsFactors = FALSE)
  US_Ht_CBH <- read.csv(file = paste0(path, s_s_prefix, table_folder, "US_Ht_CBH.csv"), row.names = NULL, stringsAsFactors = FALSE)
#
coniferList <- c("Ba","Bl", "Bg","Cw", "Fd", "Hw","T","Pl", "Sx" , "Lw", "Lt")
nonConiferList <- c("Act","Ac","At","Ep","DP", "DU","Dead","DP")

  # Remove columns that start with NA
  US_SPH <- US_SPH[, !grepl("^NA", names(US_SPH))]
  
  # Check if there is any overstory data
  has_overstory <- sum(OS_SPH[2]) > 0

  if(has_overstory){
    # 1. Overstorey
    
    # Height and CBH
    OS_Ht_CBH_melt <- melt(OS_Ht_CBH, id = "DBH.Class")
    OS_Ht_CBH_melt$Species <- substr(OS_Ht_CBH_melt$variable, start = 1, stop = 2)
    
    OS_Ht <- OS_Ht_CBH_melt[grep(pattern = "Ht", OS_Ht_CBH_melt$variable, fixed = TRUE), ]
    names(OS_Ht)[names(OS_Ht) == "value"] <- "Mean_Height"
    OS_Ht <- subset(OS_Ht, select = -c(variable))
    
    OS_CBH <- OS_Ht_CBH_melt[grep(pattern = "CBH", OS_Ht_CBH_melt$variable, fixed = TRUE), ]
    names(OS_CBH)[names(OS_CBH) == "value"] <- "Mean_CBH"
    OS_CBH <- subset(OS_CBH, select = -c(variable))
    
    OS_output <- merge(OS_CBH, OS_Ht, all = TRUE)
    
    # Species per hectare
    OS_SPH_melt <- melt(OS_SPH, id = "DBH.Class")
    setnames(OS_SPH_melt, old = c("variable", "value"), new = c("Species", "SPH"))
    
    OS_output <- merge(OS_output, OS_SPH_melt, by = c("Species", "DBH.Class"), all = TRUE)
    
    # Basal area
    OS_BA_melt <- melt(OS_BA, id = "DBH.Class")
    setnames(OS_BA_melt, old = c("variable", "value"), new = c("Species", "BA"))
    
    OS_output <- merge(OS_output, OS_BA_melt, by = c("Species", "DBH.Class"), all = TRUE)
    
    # Volume
    OS_Vol_melt <- melt(OS_Vol, id = "DBH.Class")
    setnames(OS_Vol_melt, old = c("variable", "value"), new = c("Species", "Volume"))
    
    OS_output <- merge(OS_output, OS_Vol_melt, by = c("Species", "DBH.Class"), all = TRUE)
    
    # Rename DBH class
    setnames(OS_output, old = c("DBH.Class"), new = c("DBH_class"))
    
    # Replace DBH class names with layer descriptions
    OS_output[OS_output == "12.5 - 17.5"] <- "Layer 1 (12.5 - 17.5 cm dbh)"
    OS_output[OS_output == "17.5 - 22.5"] <- "Layer 1 (17.5 - 22.5 cm dbh)"
    OS_output[OS_output == "22.5 - 27.5"] <- "Layer 1 (22.5 - 27.5 cm dbh)"
    OS_output[OS_output == "27.5 - 35"] <- "Layer 1 (27.5 - 35 cm dbh)"
    OS_output[OS_output == "35 - 45"] <- "Layer 1 (35 - 45 cm dbh)"
    OS_output[OS_output == "45+"] <- "Layer 1 (45+ cm dbh)"
  } else {
    OS_output <- NULL
  }

  # 2. Understorey
  # Height and CBH
  US_Ht_CBH_melt <- melt(US_Ht_CBH, id = "Layer")
  US_Ht_CBH_melt$Species <- substr(US_Ht_CBH_melt$variable, start = 1, stop = 2)
  US_Ht_CBH_melt <- US_Ht_CBH_melt[US_Ht_CBH_melt$Species != "De", ]

  US_Ht <- US_Ht_CBH_melt[grep(pattern = "Ht", US_Ht_CBH_melt$variable, fixed = TRUE), ]
  names(US_Ht)[names(US_Ht) == "value"] <- "Mean_Height"
  US_Ht <- subset(US_Ht, select = -c(variable))
  
  US_CBH <- US_Ht_CBH_melt[grep(pattern = "CBH", US_Ht_CBH_melt$variable, fixed = TRUE), ]
  names(US_CBH)[names(US_CBH) == "value"] <- "Mean_CBH"
  US_CBH <- subset(US_CBH, select = -c(variable))
  
  US_output <- merge(US_CBH, US_Ht, all = TRUE)
  
  # Species per hectare
  US_SPH_melt <- melt(US_SPH, id = "Layer")
  setnames(US_SPH_melt, old = c("variable", "value"), new = c("Species", "SPH"))
  
  US_output <- merge(US_output, US_SPH_melt, by = c("Species", "Layer"), all = TRUE)
  
  # Add placeholder columns for Basal area and Volume
  US_output$BA <- NA
  US_output$Volume <- NA
  
  # Rename Layer to DBH class
  setnames(US_output, old = c("Layer"), new = c("DBH_class"))
  US_output[US_output == "Layer 1 (12.5-17.5)"] <- "Layer 1 (12.5 - 17.5 cm dbh)"
  
  # Joining overstorey and understorey
  if (!is.null(OS_output)) {
    All_output <- rbind(OS_output, US_output)
  } else {
    All_output <- US_output
  }
  
  All_output <- All_output[!duplicated(All_output), ]
  # Remove totals for species across all diameter classes
  All_output <- All_output[which(All_output$DBH_class != "Total"), ]
 
  # Joining the cutting specs
  cuttingSpecs <- subset(cuttingSpecs, select = -c(Stand.Layer))
  cuttingSpecs_melt <- melt(cuttingSpecs, id = "DBH.Class")
  setnames(cuttingSpecs_melt, old = c("DBH.Class", "variable", "value"), new = c("DBH_class", "Species", "Cutting_Spec"))

  # Replace DBH class names in cutting specs
  cuttingSpecs_melt[cuttingSpecs_melt == "0-1.5"] <- "Layer 4 (<1.3m)"
  cuttingSpecs_melt[cuttingSpecs_melt == "1.5-7.5"] <- "Layer 3 (2.5-7.49)"
  cuttingSpecs_melt[cuttingSpecs_melt == "7.5-12.5"] <- "Layer 2 (7.5-12.49)"
  cuttingSpecs_melt[cuttingSpecs_melt == "12.5-17.5"] <- "Layer 1 (12.5 - 17.5 cm dbh)"
  cuttingSpecs_melt[cuttingSpecs_melt == "17.5-22.5"] <- "Layer 1 (17.5 - 22.5 cm dbh)"
  cuttingSpecs_melt[cuttingSpecs_melt == "22.5-27.5"] <- "Layer 1 (22.5 - 27.5 cm dbh)"
  cuttingSpecs_melt[cuttingSpecs_melt == "27.5-35"] <- "Layer 1 (27.5 - 35 cm dbh)"
  cuttingSpecs_melt[cuttingSpecs_melt == "35-45"] <- "Layer 1 (35 - 45 cm dbh)"
  cuttingSpecs_melt[cuttingSpecs_melt == "45+"] <- "Layer 1 (45+ cm dbh)"
  
  All_output <- merge(All_output, cuttingSpecs_melt, by = c("Species", "DBH_class"), all.x = TRUE)
  All_output <- All_output[All_output$Species != "Dead", ]

  # Calculating cut and leave
  All_output <- All_output %>%
    mutate(SPH_cut = round(SPH * Cutting_Spec / 100, digits = 0)) %>%
    mutate(SPH_leave = round(SPH * (100 - Cutting_Spec) / 100, digits = 0)) %>%
    mutate(Volume_cut = round(Volume * Cutting_Spec / 100, digits = 0)) %>%
    mutate(Volume_leave = round(Volume * (100 - Cutting_Spec) / 100, digits = 0)) %>%
    mutate(BA_cut = round(BA * Cutting_Spec / 100, digits = 0)) %>%
    mutate(BA_leave = round(BA * (100 - Cutting_Spec) / 100, digits = 0))
  
  All_output <- All_output[!is.na(All_output$SPH), ]
  All_output <- All_output[All_output$SPH != 0, ]
  All_output <- All_output[!All_output$Species == "Total", ]
  All_output[is.na(All_output)] <- 0

  All_output <- All_output %>%
    filter(!is.na(Species))

  # Set height values for dead trees
  All_output[which(All_output$Species == "DP" & All_output$DBH_class == "Layer 1 (12.5 - 17.5 cm dbh)"), c("Mean_Height")] <- 8
  All_output[which(All_output$Species == "DU" & All_output$DBH_class == "Layer 3 (2.5-7.49)"), c("Mean_Height")] <- 2
  All_output[which(All_output$Species == "DU" & All_output$DBH_class == "Layer 4 (<1.3m)"), c("Mean_Height")] <- 1
`
  # Calculating Totals
  All_output <- mutate(All_output, Decay = ifelse(All_output$Species == "DP" | All_output$Species == "DU", "Dead", "Live"))
  All_output <- mutate(All_output, Conifer = ifelse(All_output$Species %in% coniferList, "Coniferous", "Other"))
  
  All_output$Mean_CBH[All_output$Conifer != "Coniferous"] <- NA
  All_output$Mean_CBH <- as.integer(All_output$Mean_CBH)
  All_output$Mean_Height <- as.integer(All_output$Mean_Height)
  
  All_output <- mutate(All_output, CBH_forWtMean = All_output$Mean_CBH * All_output$SPH)
  All_output <- mutate(All_output, Ht_forWtMean = All_output$Mean_Height * All_output$SPH)

  # Conifer
  coniferTemp <- aggregate(cbind(SPH, SPH_cut, SPH_leave, Volume, Volume_cut, Volume_leave,
                                 BA, BA_cut, BA_leave, CBH_forWtMean, Ht_forWtMean) ~ Conifer + DBH_class,
                           data = All_output, FUN = sum, na.action = na.pass)
  
  coniferTemp <- coniferTemp %>%
    filter(Conifer == "Coniferous") %>%
    mutate(Species = "Total Conifer") %>%
    mutate(Mean_CBH = round(CBH_forWtMean / SPH, digits = 1)) %>%
    mutate(Mean_Height = round(Ht_forWtMean / SPH, digits = 1)) %>%
    dplyr::select(-Conifer, -CBH_forWtMean, -Ht_forWtMean)
  
  existingLayers <- c("Layer 4 (<1.3m)",
                      "Layer 3 (2.5-7.49)",
                      "Layer 2 (7.5-12.49)",
                      "Layer 1 (12.5 - 17.5 cm dbh)",
                      "Layer 1 (17.5 - 22.5 cm dbh)",
                      "Layer 1 (22.5 - 27.5 cm dbh)",
                      "Layer 1 (27.5 - 35 cm dbh)",
                      "Layer 1 (35 - 45 cm dbh)",
                      "Layer 1 (45+ cm dbh)")
  
  coniferLayers <- unique(coniferTemp$DBH_class)
  missingLayers <- existingLayers[!(existingLayers %in% coniferLayers)]
  
  if (length(missingLayers) > 0) {
    coniferTempFill <- data.frame(Species = "Total Conifer", DBH_class = missingLayers,
                                  SPH = 0, SPH_cut = 0, SPH_leave = 0,
                                  Volume = 0, Volume_cut = 0, Volume_leave = 0,
                                  BA = 0, BA_cut = 0, BA_leave = 0,
                                  Mean_CBH = NA, Mean_Height = NA)
    
    coniferTemp <- rbind.fill(coniferTemp, coniferTempFill)
  }
  
  # All Live
  liveTemp <- aggregate(cbind(SPH, SPH_cut, SPH_leave,
                              Volume, Volume_cut, Volume_leave,
                              BA, BA_cut, BA_leave, CBH_forWtMean, Ht_forWtMean) ~ Decay + DBH_class,
                        data = All_output, FUN = sum, na.action = na.pass)
  
  liveTemp <- liveTemp %>%
    filter(Decay == "Live") %>%
    mutate(Species = "Total Live") %>%
    mutate(Mean_CBH = round(CBH_forWtMean / SPH, digits = 1)) %>%
    mutate(Mean_Height = round(Ht_forWtMean / SPH, digits = 1)) %>%
    dplyr::select(-Decay, -CBH_forWtMean, -Ht_forWtMean)
  
  liveLayers <- unique(liveTemp$DBH_class)
  missingLayers <- existingLayers[!(existingLayers %in% liveLayers)]
  
  if (length(missingLayers) > 0) {
    liveTempFill <- data.frame(Species = "Total Live", DBH_class = missingLayers,
                               SPH = 0, SPH_cut = 0, SPH_leave = 0,
                               Volume = 0, Volume_cut = 0, Volume_leave = 0,
                               BA = 0, BA_cut = 0, BA_leave = 0,
                               Mean_CBH = NA, Mean_Height = NA)
    
    liveTemp <- rbind.fill(liveTemp, liveTempFill)
  }
  
  # Everything per layer
  allPerLayerTemp <- aggregate(cbind(SPH, SPH_cut, SPH_leave,
                                     Volume, Volume_cut, Volume_leave,
                                     BA, BA_cut, BA_leave, CBH_forWtMean, Ht_forWtMean) ~ DBH_class,
                               data = All_output, FUN = sum, na.action = na.pass)
  
  allPerLayerTemp <- allPerLayerTemp %>%
    mutate(Species = "Total All Species") %>%
    mutate(Mean_CBH = round(CBH_forWtMean / SPH, digits = 1)) %>%
    mutate(Mean_Height = round(Ht_forWtMean / SPH, digits = 1)) %>%
    dplyr::select(-CBH_forWtMean, -Ht_forWtMean)
  
  allLayers <- unique(allPerLayerTemp$DBH_class)
  missingLayers <- existingLayers[!(existingLayers %in% allLayers)]
  
  if (length(missingLayers) > 0) {
    allPerLayerTempFill <- data.frame(Species = "Total All Species", DBH_class = missingLayers,
                                      SPH = 0, SPH_cut = 0, SPH_leave = 0,
                                      Volume = 0, Volume_cut = 0, Volume_leave = 0,
                                      BA = 0, BA_cut = 0, BA_leave = 0,
                                      Mean_CBH = NA, Mean_Height = NA)
    
    allPerLayerTemp <- rbind.fill(allPerLayerTemp, allPerLayerTempFill)
  }
  
  # Everything per layer
  allTemp <- data.frame(Species = "Total All Layers", DBH_class = NA, SPH = sum(All_output$SPH, na.rm = T), SPH_cut = sum(All_output$SPH_cut, na.rm = T), SPH_leave = sum(All_output$SPH_leave, na.rm = T), Volume = sum(All_output$Volume, na.rm = T), Volume_cut = sum(All_output$Volume_cut, na.rm = T), Volume_leave = sum(All_output$Volume_leave, na.rm = T), BA = sum(All_output$BA, na.rm = T), BA_cut = sum(All_output$BA_cut, na.rm = T), BA_leave = sum(All_output$BA_leave, na.rm = T), Mean_CBH = NA, Ht_forWtMean = sum(All_output$Ht_forWtMean, na.rm = T))
  
  allTemp <- allTemp %>%
    mutate(Mean_Height = round(Ht_forWtMean / SPH, digits = 1)) %>%
    dplyr::select(-c("Ht_forWtMean"))


  # Layer-1 specific
  if (!is.null(OS_output)) {
    L1temp <- All_output[grep(pattern = "Layer 1", All_output$DBH_class, fixed = TRUE), ]
   
    # Conifers
    L1conifer <- aggregate(cbind(SPH, SPH_cut, SPH_leave,
                                 Volume, Volume_cut, Volume_leave,
                                 BA, BA_cut, BA_leave, CBH_forWtMean, Ht_forWtMean) ~ Conifer,
                           data = L1temp, FUN = sum, na.rm = TRUE, na.action = na.pass)
    
    L1conifer <- L1conifer %>%
      filter(Conifer == "Coniferous") %>%
      mutate(Species = "Total Layer 1 - Conifers Only") %>%
      mutate(DBH_class = "Total Layer 1") %>%
      mutate(Mean_CBH = round(CBH_forWtMean / SPH, digits = 1)) %>%
      mutate(Mean_Height = round(Ht_forWtMean / SPH, digits = 1)) %>%
      dplyr::select(-c("Conifer", "CBH_forWtMean", "Ht_forWtMean"))
    
    # All Species
    L1all <- L1temp %>%
      dplyr::select(SPH, SPH_cut, SPH_leave,
                    Volume, Volume_cut, Volume_leave,
                    BA, BA_cut, BA_leave, CBH_forWtMean, Ht_forWtMean)
    
    L1all <- as.data.frame(t(colSums(L1all, na.rm = TRUE)))
    
    L1all <- L1all %>%
      mutate(Species = "Total Layer 1 - All Species") %>%
      mutate(DBH_class = "Total Layer 1") %>%
      mutate(Mean_CBH = round(CBH_forWtMean / SPH, digits = 1)) %>%
      mutate(Mean_Height = round(Ht_forWtMean / SPH, digits = 1)) %>%
      dplyr::select(-c("CBH_forWtMean", "Ht_forWtMean"))
    
    # Dead Potential
    existingDPlayers <- All_output$DBH_class[All_output$Species == "DP"]
    
    allDPlayers <- c("Layer 2 (7.5-12.49)",
                     "Layer 1 (12.5 - 17.5 cm dbh)",
                     "Layer 1 (17.5 - 22.5 cm dbh)",
                     "Layer 1 (22.5 - 27.5 cm dbh)",
                     "Layer 1 (27.5 - 35 cm dbh)",
                     "Layer 1 (35 - 45 cm dbh)",
                     "Layer 1 (45+ cm dbh)")
    
    missingDPlayers <- allDPlayers[!(allDPlayers %in% existingDPlayers)]
    
    DPtemp <- data.frame(Species = "DP", DBH_class = missingDPlayers,
                         SPH = 0, SPH_cut = 0, SPH_leave = 0,
                         Volume = 0, Volume_cut = 0, Volume_leave = 0,
                         BA = 0, BA_cut = 0, BA_leave = 0)
    
    # Put everything together...
    All_output <- rbind.fill(All_output, coniferTemp, liveTemp, allPerLayerTemp, allTemp, L1conifer, L1all, DPtemp)
  } else {
    # No overstory, just add the understory and totals
    All_output <- rbind.fill(All_output, coniferTemp, liveTemp, allPerLayerTemp, allTemp)
  }
  
  # Reformatting
  All_output <- All_output[, c("Species", "DBH_class", "Mean_CBH", "Mean_Height",
                               "SPH", "SPH_cut", "SPH_leave",
                               "Volume", "Volume_cut", "Volume_leave",
                               "BA", "BA_cut", "BA_leave",
                               "Cutting_Spec")]
  
  setnames(All_output, old = c("SPH", "Volume", "BA", "Cutting_Spec"), new = c("SPH_exist", "Volume_exist", "BA_exist", "Cutting_spec"))

  All_output$Species[All_output$Species == "DP"] <- "Total Dead Potential"
  All_output$DBH_class[All_output$DBH_class == "Layer 2 (7.5-12.49)"] <- "Layer 2 (7.5 - 12.49 cm dbh)"
  All_output$DBH_class[All_output$DBH_class == "Layer 3 (2.5-7.49)"] <- "Layer 3 (2.5 - 7.49 cm dbh)"
  All_output$DBH_class[All_output$DBH_class == "Layer 4 (<1.3m)"] <- "Layer 4 (<1.3 cm dbh)"
  
  All_output$Cutting_spec[is.na(All_output$Cutting_spec)] <- 0
  
  totallingLayers <- c("Total All Species",
                       "Total Dead Potential",
                       "Total Live",
                       "Total Layer 1 - All Species")
  
  VolBA <- c("Volume_exist",
             "Volume_cut",
             "Volume_leave",
             "BA_exist",
             "BA_cut",
             "BA_leave")
  
  USlayers <- c("Layer 2 (7.5 - 12.49 cm dbh)",
                "Layer 3 (2.5 - 7.49 cm dbh)",
                "Layer 4 (<1.3 cm dbh)")
  
  All_output$Mean_CBH[which(All_output$Species %in% totallingLayers)] <- "-"
  All_output[which(All_output$DBH_class %in% USlayers), which(colnames(All_output) %in% VolBA)] <- "-"
  All_output[is.na(All_output)] <- "-"
  
  All_output$Species <- factor(All_output$Species, levels = c(coniferList,
                                                              nonConiferList[!nonConiferList == "DP"],
                                                              "Total Dead Potential",
                                                              "Total Live",
                                                              "Total All Species",
                                                              "Total Layer 1 - All Species",
                                                              "Total Conifer",
                                                              "Total Layer 1 - Conifers Only"), ordered = TRUE)
  
  All_output$DBH_class <- factor(All_output$DBH_class, levels = c("Layer 1 (45+ cm dbh)",
                                                                  "Layer 1 (35 - 45 cm dbh)",
                                                                  "Layer 1 (27.5 - 35 cm dbh)",
                                                                  "Layer 1 (22.5 - 27.5 cm dbh)",
                                                                  "Layer 1 (17.5 - 22.5 cm dbh)",
                                                                  "Layer 1 (12.5 - 17.5 cm dbh)",
                                                                  "Total Layer 1",
                                                                  "Layer 2 (7.5 - 12.49 cm dbh)",
                                                                  "Layer 3 (2.5 - 7.49 cm dbh)",
                                                                  "Layer 4 (<1.3 cm dbh)"), ordered = TRUE)


  All_output <- All_output[with(All_output, order(DBH_class, Species)), ]
  All_output$Species <- as.character(All_output$Species)
  All_output$DBH_class <- as.character(All_output$DBH_class)
  
  # Extract layer totals
  allTemp <- tail(All_output, n = 1)
  All_output <- slice(All_output, 1:(n() - 1))
  
  allTemp$Species <- "TOTAL ALL LAYERS"
  allTemp$DBH_class <- "Layer Totals"
  
  USlayers <- c("Layer 2 (7.5 - 12.49 cm dbh)",
                "Layer 3 (2.5 - 7.49 cm dbh)",
                "Layer 4 (<1.3 cm dbh)")
  
  totalsL1 <- All_output[All_output$Species == "Total Layer 1 - All Species" & All_output$DBH_class == "Total Layer 1", ]
  totalsUS <- All_output[All_output$Species == "Total All Species" & All_output$DBH_class %in% USlayers, ]
  
  totals <- rbind.fill(totalsL1, totalsUS, allTemp)
  totals <- totals[grepl("^\\d+\\.?\\d*$", totals$SPH_exist), ]
  

  #modifying if no overstory
  species_totals_labels<- c("Total Layer 1",
    "Layer 2 (7.5 - 12.49 cm dbh)",
    "Layer 3 (2.5 - 7.49 cm dbh)",
    "Layer 4 (<1.3 cm dbh)",
    "Layer Totals")


# Check if all elements of species_totals_labels are in totals$DBH_class
  if(all(species_totals_labels %in% totals$DBH_class)){
  # Format and append totals
  totals$Species <- c("Total All Species Layer 1",
                      "Total All Species Layer 2",
                      "Total All Species Layer 3",
                      "Total All Species Layer 4",
                      "TOTAL ALL LAYERS")
} else {
  layer_1df<- data.frame(Species="Total Layer 1 - All Species",DBH_class="Total Layer 1", Mean_CBH="-",Mean_Height="-",SPH_exist=0,SPH_cut=0,SPH_leave=0,Volume_exist=0,Volume_cut=0,Volume_leave=0,BA_exist=0,BA_cut=0,BA_leave=0,Cutting_spec=0)
  totals<- rbind(layer_1df,totals)
  totals$Species <- c("Total All Species Layer 1",
                      "Total All Species Layer 2",
                      "Total All Species Layer 3",
                      "Total All Species Layer 4",
                      "TOTAL ALL LAYERS")
}
  totals$DBH_class <- "Layer Totals"
  All_output <- rbind.fill(All_output, totals)

  # Write the fused table as a CSV
  write.csv(All_output, file = paste0(path, s_s_prefix, table_folder, TR, "_SST.csv"), row.names = F)
  
  #####################################################################################################################

    data.stand <- read.csv(paste0(path, s_s_prefix, table_folder, TR, "_SST.csv"), header = TRUE)
    # Update start & end row number for the output_table code block (starting at line 19) by viewing the data.stand, 
    value_exists <- any(data.stand$DBH_class == "Total Layer 1")
    
if(isTRUE(value_exists)){    
  
    ###then run the following code###
    data.stand.final <- dplyr::select(data.stand, -c(DBH_class, Cutting_spec))
    colnames(data.stand.final) <- c("Species", "Average Crown to Base Height (m)",
                                    "Average Tree Height (m)", "Existing", "Cut",
                                    "Leave", "Existing", "Cut",
                                    "Leave", "Existing", "Cut",
                                    "Leave")


    ########################start formatting the output table###########################################################################
    # Please note that the start and end row number for each layer need to be updated before running the code below
    output_table <- kbl(data.stand.final, caption = "", align = "c") %>%
      kable_paper(full_width = F) %>%
      pack_rows("Layer 1 (>45 cm dbh)", 
                which(data.stand$DBH_class == "Layer 1 (45+ cm dbh)")[1],
                (which(data.stand$DBH_class == "Layer 1 (45+ cm dbh)")[(length(which(data.stand$DBH_class == "Layer 1 (45+ cm dbh)")))]),
                label_row_css = "background-color: #777; color: #fff;") %>%
      pack_rows("Layer 1 (35 - 45 cm dbh)",
                which(data.stand$DBH_class == "Layer 1 (35 - 45 cm dbh)")[1],
                (which(data.stand$DBH_class == "Layer 1 (35 - 45 cm dbh)")[(length(which(data.stand$DBH_class == "Layer 1 (35 - 45 cm dbh)")))]),
                label_row_css = "background-color: #777; color: #fff;") %>%
      pack_rows("Layer 1 (27.5 - 35 cm dbh)",
                which(data.stand$DBH_class == "Layer 1 (27.5 - 35 cm dbh)")[1],
                (which(data.stand$DBH_class == "Layer 1 (27.5 - 35 cm dbh)")[(length(which(data.stand$DBH_class == "Layer 1 (27.5 - 35 cm dbh)")))]),
                label_row_css = "background-color: #777; color: #fff;") %>%
      pack_rows("Layer 1 (22.5 - 27.5 cm dbh)", 
                which(data.stand$DBH_class == "Layer 1 (22.5 - 27.5 cm dbh)")[1],
                (which(data.stand$DBH_class == "Layer 1 (22.5 - 27.5 cm dbh)")[(length(which(data.stand$DBH_class == "Layer 1 (22.5 - 27.5 cm dbh)")))]),
                label_row_css = "background-color: #777; color: #fff;") %>%
      pack_rows("Layer 1 (17.5 - 22.5 cm dbh)",
                which(data.stand$DBH_class == "Layer 1 (17.5 - 22.5 cm dbh)")[1],
                (which(data.stand$DBH_class == "Layer 1 (17.5 - 22.5 cm dbh)")[(length(which(data.stand$DBH_class == "Layer 1 (17.5 - 22.5 cm dbh)")))]),
                label_row_css = "background-color: #777; color: #fff;") %>%
      pack_rows("Layer 1 (12.5 - 17.5 cm dbh)", 
                which(data.stand$DBH_class == "Layer 1 (12.5 - 17.5 cm dbh)")[1],
                (which(data.stand$DBH_class == "Layer 1 (12.5 - 17.5 cm dbh)")[(length(which(data.stand$DBH_class == "Layer 1 (12.5 - 17.5 cm dbh)")))]),
                label_row_css = "background-color: #777; color: #fff;") %>%
      pack_rows("Total Layer 1", 
                which(data.stand$DBH_class == "Total Layer 1")[1],
                (which(data.stand$DBH_class == "Total Layer 1")[(length(which(data.stand$DBH_class == "Total Layer 1")))]),
                label_row_css = "background-color: #777; color: #fff;") %>%
      pack_rows("Layer 2 (7.5 - 12.49 cm dbh)", 
                which(data.stand$DBH_class == "Layer 2 (7.5 - 12.49 cm dbh)")[1],
                (which(data.stand$DBH_class == "Layer 2 (7.5 - 12.49 cm dbh)")[(length(which(data.stand$DBH_class == "Layer 2 (7.5 - 12.49 cm dbh)")))]),
                label_row_css = "background-color: #777; color: #fff;") %>%
      pack_rows("Layer 3 (2.5 - 7.49 cm dbh)",
                which(data.stand$DBH_class == "Layer 3 (2.5 - 7.49 cm dbh)")[1],
                (which(data.stand$DBH_class == "Layer 3 (2.5 - 7.49 cm dbh)")[(length(which(data.stand$DBH_class == "Layer 3 (2.5 - 7.49 cm dbh)")))]),
                label_row_css = "background-color: #777; color: #fff;") %>%
      pack_rows("Layer 4 (<1.3 cm dbh)",
                which(data.stand$DBH_class == "Layer 4 (<1.3 cm dbh)")[1],
                (which(data.stand$DBH_class == "Layer 4 (<1.3 cm dbh)")[(length(which(data.stand$DBH_class == "Layer 4 (<1.3 cm dbh)")))]),
                label_row_css = "background-color: #777; color: #fff;") %>%
      pack_rows("Layer Totals",
                which(data.stand$DBH_class == "Layer Totals")[1],
                (which(data.stand$DBH_class == "Layer Totals")[(length(which(data.stand$DBH_class == "Layer Totals")))]),
                label_row_css = "background-color: #777; color: #fff;") %>%
      add_header_above(c("","","", "STEM PER HECTARE (SPH)" = 3, "VOLUME (m3/ha)" = 3, "BASAL AREA (m2/ha)" = 3)) %>%
      add_header_above(c("STAND AND STOCK TABLE " = 12),
                       align = "left")
    
    save_kable(output_table, file = paste0(path, s_s_prefix, table_folder, TR, "_SST.html"))
   }
else{
    ###then run the following code###
    data.stand.final <- dplyr::select(data.stand, -c(DBH_class, Cutting_spec))
    colnames(data.stand.final) <- c("Species", "Average Crown to Base Height (m)",
                                    "Average Tree Height (m)", "Existing", "Cut",
                                    "Leave", "Existing", "Cut",
                                    "Leave", "Existing", "Cut",
                                    "Leave")


    ########################start formatting the output table###########################################################################
    # Please note that the start and end row number for each layer need to be updated before running the code below
    output_table <- kbl(data.stand.final, caption = "", align = "c") %>%
      kable_paper(full_width = F) %>%
      pack_rows("Layer 1 (>45 cm dbh)", 
                which(data.stand$DBH_class == "Layer 1 (45+ cm dbh)")[1],
                (which(data.stand$DBH_class == "Layer 1 (45+ cm dbh)")[(length(which(data.stand$DBH_class == "Layer 1 (45+ cm dbh)")))]),
                label_row_css = "background-color: #777; color: #fff;") %>%
      pack_rows("Layer 1 (35 - 45 cm dbh)",
                which(data.stand$DBH_class == "Layer 1 (35 - 45 cm dbh)")[1],
                (which(data.stand$DBH_class == "Layer 1 (35 - 45 cm dbh)")[(length(which(data.stand$DBH_class == "Layer 1 (35 - 45 cm dbh)")))]),
                label_row_css = "background-color: #777; color: #fff;") %>%
      pack_rows("Layer 1 (27.5 - 35 cm dbh)",
                which(data.stand$DBH_class == "Layer 1 (27.5 - 35 cm dbh)")[1],
                (which(data.stand$DBH_class == "Layer 1 (27.5 - 35 cm dbh)")[(length(which(data.stand$DBH_class == "Layer 1 (27.5 - 35 cm dbh)")))]),
                label_row_css = "background-color: #777; color: #fff;") %>%
      pack_rows("Layer 1 (22.5 - 27.5 cm dbh)", 
                which(data.stand$DBH_class == "Layer 1 (22.5 - 27.5 cm dbh)")[1],
                (which(data.stand$DBH_class == "Layer 1 (22.5 - 27.5 cm dbh)")[(length(which(data.stand$DBH_class == "Layer 1 (22.5 - 27.5 cm dbh)")))]),
                label_row_css = "background-color: #777; color: #fff;") %>%
      pack_rows("Layer 1 (17.5 - 22.5 cm dbh)",
                which(data.stand$DBH_class == "Layer 1 (17.5 - 22.5 cm dbh)")[1],
                (which(data.stand$DBH_class == "Layer 1 (17.5 - 22.5 cm dbh)")[(length(which(data.stand$DBH_class == "Layer 1 (17.5 - 22.5 cm dbh)")))]),
                label_row_css = "background-color: #777; color: #fff;") %>%
      pack_rows("Layer 1 (12.5 - 17.5 cm dbh)", 
                which(data.stand$DBH_class == "Layer 1 (12.5 - 17.5 cm dbh)")[1],
                (which(data.stand$DBH_class == "Layer 1 (12.5 - 17.5 cm dbh)")[(length(which(data.stand$DBH_class == "Layer 1 (12.5 - 17.5 cm dbh)")))]),
                label_row_css = "background-color: #777; color: #fff;") %>%
      pack_rows("Layer 2 (7.5 - 12.49 cm dbh)", 
                which(data.stand$DBH_class == "Layer 2 (7.5 - 12.49 cm dbh)")[1],
                (which(data.stand$DBH_class == "Layer 2 (7.5 - 12.49 cm dbh)")[(length(which(data.stand$DBH_class == "Layer 2 (7.5 - 12.49 cm dbh)")))]),
                label_row_css = "background-color: #777; color: #fff;") %>%
      pack_rows("Layer 3 (2.5 - 7.49 cm dbh)",
                which(data.stand$DBH_class == "Layer 3 (2.5 - 7.49 cm dbh)")[1],
                (which(data.stand$DBH_class == "Layer 3 (2.5 - 7.49 cm dbh)")[(length(which(data.stand$DBH_class == "Layer 3 (2.5 - 7.49 cm dbh)")))]),
                label_row_css = "background-color: #777; color: #fff;") %>%
      pack_rows("Layer 4 (<1.3 cm dbh)",
                which(data.stand$DBH_class == "Layer 4 (<1.3 cm dbh)")[1],
                (which(data.stand$DBH_class == "Layer 4 (<1.3 cm dbh)")[(length(which(data.stand$DBH_class == "Layer 4 (<1.3 cm dbh)")))]),
                label_row_css = "background-color: #777; color: #fff;") %>%
      pack_rows("Layer Totals",
                which(data.stand$DBH_class == "Layer Totals")[1],
                (which(data.stand$DBH_class == "Layer Totals")[(length(which(data.stand$DBH_class == "Layer Totals")))]),
                label_row_css = "background-color: #777; color: #fff;") %>%
      add_header_above(c("","","", "STEM PER HECTARE (SPH)" = 3, "VOLUME (m3/ha)" = 3, "BASAL AREA (m2/ha)" = 3)) %>%
      add_header_above(c("STAND AND STOCK TABLE " = 12),
                       align = "left")
    
    save_kable(output_table, file = paste0(path, s_s_prefix, table_folder, TR, "_SST.html"))
    }

#}

```